### 2022-02-14 Msc_Thesis_Lea_v2.Rmd
Create Msc_Thesis_Lea_v2.Rmd

---
title: "Cross-sectoral conservation of the Indian Ocean"
author: "Lea Fourchault"
date: "02/02/2022"
output: html_document
---
<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 34px;
  color: Black;
}
h1 { /* Header 1 */
  font-size: 24px;
  color: Black;
}
h2 { /* Header 2 */
    font-size: 20px;
  color: Black;
}
h3 { /* Header 3 */
  font-size: 14px;
  color: Black;
}
code.r{ /* Code block */
    font-size: 11px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Table of Contents
# Introduction
# Materials & Methods
## Workflow with Prioritizr 
## Overview of data  & layers
## Planning region 
## Conservation features & targets
### IBAs
### IMMAs
### EBSAs
### Deep-sea features : seamounts, ridges, plateaus, vents 
## Cost layer
### CFC
### PMN
### Exploration areas
### Reserved areas
### Shipping
### Fishing *use cost layer already created?*
# Prioritizr problems
### Basic run (only Cons features)
### Cons features, fishing
### Cons features, mining (lock-out expl  & res areas; assign cost for PMN areas)
### Cons features, fishing + mining 
### Cons features, shipping 
# Discussion
## Potential improvements
# Conclusion

---

Current questions:

+ how to group all files (ex: all EBSAs + IBAs + IMMAs + vents + seamounts...) into one, to create ConsFeatures?
+ should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection
+ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what coupd be other methods (not using as.polygon)?
+ shipping raster : cells have certain values (# of ships recorded??I will check for Tuesday), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)? So we wouldn't try to compute the cost of deviating from any given route, but rather compute the price for each ship to go through 1 PU * the number of ships going through that PU. It sounds fairly logical to me, but maybe it's not?
+ fishing cost layer : there is a ready-made fx in SpatPlan repo (thanks Jase & Tin), should we use it for this project? see fishing subsection (but issues with plotting  as you can see)

# Introduction

This project aims to identify areas for MPAs that maximise the conservation of biologically and ecologically important areas in the Indian Ocean's ABNJS, across all depths, while minimizing conflict with local industrial stakeholders - including the fishing, shipping, and deep-sea mining industries.

Why ABNJs?

+ to reach the global target of 30% marine protection, ABNJs are optimal (largest area; common good: higher incentive to protect since lower opportunity cost for individual countries)

Why the Indian Ocean?

+ numerous ecosystems providing important services and unique biodiversity, but few already existing MPAs within EEZs and none in ABNJs
+ active exploration of PMN & CFC for deep sea mining but no plan for conservation linked to deep-sea mining (unlike Pacific)
+ many stakeholders present (increasingly dense shipping routes, fishing activity, deep-sea mining): higher risk of conflict: prioritization approach should be useful

Why considering all depth layers?

+ despite unique biodiversity, the deep sea is often ignored in MPA design
+ the deep sea increasingly at risk from anthropogenic disturbances
+ we need to consider all layers to ensure vertical connectivity
+ the implementation of MPAs is only feasible on surface level (makes no sense to focus only on deep-sea, even if major preoccupation is benthic or at high depth).

Why considering multiple stakeholders?

+ so far, most MPA design is based on the fishing industry alone (despite some rather unsuccessful attempts to include tourism or other industries)
+ deep-sea mining has not yet been considered in MPA design by prioritization, despite major foreseen impacts on marine biodiversity
+ shipping routes (cargo, tankers, tourism) also disturb marine fauna, especially charismatic megafauna

```{r, cache = FALSE, echo = FALSE, results = "hide", message = FALSE}
#Open all the packages and get extras
options(repos = list(CRAN="http://cran.rstudio.com/")) 
library(tidyverse)
library(sf)
library(patchwork)
library(terra)
library(stars)
library(rnaturalearth)
library(rnaturalearthdata)
library(prioritizr)
library(knitr)

source("SpatPlan_Extras.R")

fSpatPlan_Get_Polyg <- function(filen, PUs){

  polyg_crs <- st_crs(PUs)
  Polyg1 <- st_read(filen) %>% 
    st_transform(polyg_crs)

  Polyg1PUs <- PUs %>% 
    st_centroid() %>% #Second, get all the PUs with < 50 % area on land #?
    st_within(Polyg1, sparse = FALSE) %>% 
    rowSums() %>%
    as.logical()

  # Polyg1PUs <- PUs %>% 
  #   mutate(locked_in = Polyg1PUs) ##assign new column to PUs (call it name of shp)
  # 
  return(Polyg1PUs)
}
```

# Materials and Methods

## Prioritizr workflow

Inputs needed:

+ planning region & PUs
+ conservation features & conservation targets
+ cost layer

## Layers overview

still need to

+ **select specific fishing rasters (bycatch/demersal...)/ use the fishing cost layer  already created in Spatial Planning repo**
+ and **select specific benthic shps (seamounts, ridges, plateaus, vents?)**
+ **need to create subset without inactive vents** *filter fx not working because object=character, will ask on Slack*

blue habitats **do not include hydro vents**, so will use Beaulieu & Szanfranski (2020) [link](https://doi.pangaea.de/10.1594/PANGAEA.917894?format=html#download)

Input type   |  Topic                 | Reference/source (format)
:-----------:|:----------------------:|:----------------------------------:
 COST        |  Expl + Res (mining)   |  ISA (shp)
 COST        |  PMN & CFC (mining)    |  Petersen et al., 2016 (shp)
 COST        |  Shipping              |  Halpern et al., 2013 (tif)
 COST        |  Fishing               |  Halpern et al., 2013 (tif)
 CONS        |  IBAs                  |  Birdlife datazone (shp)
 CONS        |  IMMAs                 |  importantmarinemammalhabitat (shp)
 CONS        |  EBSAs                 |  CBD (shp)
 CONS        |  Benthic features      |  Blue Habitats (shp)
 CONS        |  Longhurst provinces ? |  Marine Regions (shp)
 CONS        |  Vents                 |  Beaulieu & Szanfranski (2020) (csv)

## Planning region

```{r Planning Region, echo = TRUE, message = FALSE, results = "hide", warning = FALSE}

cCRS <- "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs" # Robinson projection
world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  st_transform(cCRS)
Region <- c(xmin = 18, xmax = 130, ymin = -39, ymax = 34) # Indian Ocean limits
Bndry <- fSpatPlan_Get_Boundary(Region, cCRS) # boundary

PU_size <- 1000 # in km2, but check resolution for this project?
Shape <- "Hexagon" # shape of PUs
MinDepth <- 0
MaxDepth <- 200 # but check for deep-sea?
PUs <- fSpatPlan_Get_PlanningUnits(Bndry, world, PU_size, Shape)

```

Boundary
```{r PR, echo=FALSE}
ggplot(data = Bndry) + geom_sf()
```

Planning Units
```{r PUs, echo=FALSE}
(ggPU <- fSpatPlan_PlotPUs(PUs, world)) # Plot Planning Units
```

## Conservation features & targets

Composed of

+ Important Bird Areas
+ Important Marine Mammal Areas
+ EBSAs
+ Deepsea features: vents, seamounts, ridges, plateaus? *(need to define)*

### Important Bird Areas

Conservation targets to be defined *need to write chunck to get ConsFeatures$amount*

```{r IBAs, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}

#loading file 
IBA <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/IBAs/MarineIBA_IndianOcean.shp"

#crossing shp & PUs
PUxIBA <- fSpatPlan_Get_Polyg(IBA, PUs)  
PUs <- PUs %>% 
  mutate(IBA = PUxIBA) #add column to PU table with
```

### Important Marine Mammal Areas

Conservation targets to be defined *need to write chunck to get ConsFeatures$amount*

```{r IMMAs, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}

#loading file
IMMA <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/IMMAs/iucn-imma.shp"

#crossing shp & PUs
PUxIMMA <- fSpatPlan_Get_Polyg(IMMA, PUs)  
PUs <- PUs %>% 
  mutate(IMMA = PUxIMMA)
```

### Ecologically and Biologically Significant Areas

Chosen to be:

+ within ABNJs (at least partially)
+ not redundant with IMMAs or IBAs

Targets: *need to write chunck to get ConsFeatures$amount*

+ NEIO_09 : 10-30%
+ NWIO_14 : 10-30%
+ SIO_11 : 10-30%
+ SIO_19 : >70%
+ SIO_22 : >70%
+ SIO_23 : >70%
+ SIO_30 : >70%
+ SIO_32 : >70%
+ SIO_35 : >70%
+ SIO_36 : >70%
+ SIO_37 : >70%

```{r EBSAs, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}

NEIO_09<- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/NEIO_9_EBSA/NEIO_9_EBSA.shp"
PUxNEIO_09 <- fSpatPlan_Get_Polyg(NEIO_09, PUs)  
PUs <- PUs %>% 
  mutate(NEIO_09 = PUxNEIO_09)

NWIO_14 <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/NWIO_14_EBSA/NWIO_14_EBSA.shp"
PUxNWIO_14 <- fSpatPlan_Get_Polyg(NWIO_14, PUs)  
PUs <- PUs %>% 
  mutate(NWIO_14 = PUxNWIO_14)

SIO_11 <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_11_EBSA-GIS shapefile/SIO_11_EBSA.shp"
PUxSIO_11 <- fSpatPlan_Get_Polyg(SIO_11, PUs)  
PUs <- PUs %>% 
  mutate(SIO_11 = PUxSIO_11)

SIO_19 <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_19_EBSA-GIS%20shapefile/SIO_19_EBSA.shp"
PUxSIO_19 <- fSpatPlan_Get_Polyg(SIO_19, PUs)  
PUs <- PUs %>% 
  mutate(SIO_19 = PUxSIO_19)

SIO_22 <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_22_EBSA-GIS shapefile/SIO_22_EBSA.shp"
PUxSIO_22 <- fSpatPlan_Get_Polyg(SIO_22, PUs)  
PUs <- PUs %>% 
  mutate(SIO_22 = PUxSIO_22)

SIO_23 <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_23_EBSA-GIS shapefile/SIO_23_EBSA.shp"
PUxSIO_23 <- fSpatPlan_Get_Polyg(SIO_23, PUs)  
PUs <- PUs %>% 
  mutate(SIO_23 = PUxSIO_23)

SIO_30 <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_30_EBSA-GIS shapefile/SIO_30_EBSA.shp"
PUx_SIO30 <- fSpatPlan_Get_Polyg(SIO_30, PUs)  
PUs <- PUs %>% 
  mutate(SIO_30 = PUx_SIO30)

SIO_32 <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_32/SIO_32_EBSA.shp"
PUxSIO_32 <- fSpatPlan_Get_Polyg(SIO_32, PUs)  
PUs <- PUs %>% 
  mutate(SIO_32 = PUxSIO_32)

SIO_35<- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_35_EBSA-GIS shapefile/SIO_35_EBSA.shp"
PUxSIO_35 <- fSpatPlan_Get_Polyg(SIO_35, PUs)  
PUs <- PUs %>% 
  mutate(SIO_35 = PUxSIO_35)

SIO_36 <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_36_EBSA-GIS shapefile/SIO_36_EBSA.shp"
PUxSIO_36 <- fSpatPlan_Get_Polyg(SIO_36, PUs)  
PUs <- PUs %>% 
  mutate(SIO_36 = PUxSIO_36)

SIO_37  <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_37_EBSA-GIS shapefile/SIO_37_EBSA.shp"
PUx_SIO37 <- fSpatPlan_Get_Polyg(SIO_37 , PUs)  
PUs <- PUs %>% 
  mutate(SIO_37 = PUx_SIO37)
```

### Deep-sea features

**Still unable to add legend in plots so showing the code instead, but will ask on Slack**

Seamounts, ridges, plateaus, vents

+ *need to add chunck for ridges & plateaus if we decide to add them*

+ *need to check how to 'filter' out inactive vents and to cross vents (point data) with PUs and how to add in plot*

Targets: tbd

Vents: >90%

Seamounts: >70%

Ridges/Plateaus: >30%

```{r Deep-sea features, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
Seamounts <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/BlueHabs/Seamounts.shp"
PUxSeamounts <- fSpatPlan_Get_Polyg(Seamounts, PUs)  
PUs <- PUs %>% 
  mutate(Seamounts = PUxSeamounts)

Vents <- read.csv("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/vents.csv")
VentsActive <- filter(Vents, Activity == "active, confirmed" | Activity == "active, inferred")
PUxVents <- fSpatPlan_Get_Polyg(VentsActive, PUs)

```

```{r Conservation features, echo = FALSE, message = FALSE, warning = FALSE}
world <- ne_countries(scale = "medium", returnclass = "sf")
IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)

IMMA <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/IMMAs/iucn-imma.shp")
IBA <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/IBAs/MarineIBA_IndianOcean.shp")
NEIO_09<- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/NEIO_9_EBSA/NEIO_9_EBSA.shp")
NWIO_14 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/NWIO_14_EBSA/NWIO_14_EBSA.shp")
SIO_11 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_11_EBSA-GIS shapefile/SIO_11_EBSA.shp")
SIO_19 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_19_EBSA-GIS%20shapefile/SIO_19_EBSA.shp")
SIO_22 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_22_EBSA-GIS shapefile/SIO_22_EBSA.shp")
SIO_23 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_23_EBSA-GIS shapefile/SIO_23_EBSA.shp")
SIO_30 <- st_read  ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_30_EBSA-GIS shapefile/SIO_30_EBSA.shp")
SIO_32 <- st_read  ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_32/SIO_32_EBSA.shp")
SIO_35<- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_35_EBSA-GIS shapefile/SIO_35_EBSA.shp")
SIO_36 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_36_EBSA-GIS shapefile/SIO_36_EBSA.shp")
SIO_37  <-st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_37_EBSA-GIS shapefile/SIO_37_EBSA.shp")
Seamounts <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/BlueHabs/Seamounts.shp")
```

```{r Conservation features plot, echo = TRUE, message = FALSE, warning = FALSE}

ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = IBA, size = 1, color = "green", fill = "green") + 
  geom_sf(data = IMMA, size = 1, color = "blue", fill = "blue")+ 
  geom_sf(data = NEIO_09, size = 1, color = "grey", fill = "grey")+ 
  geom_sf(data = NWIO_14, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_11, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_19, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_22, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_23, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_30, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_32, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_35, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_36, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_37, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = Seamounts, size = 1, color = "black", fill = "black")+
  ggtitle("IBA, IMMA, EBSAs, Seamounts") + 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
```

## Cost layer

Composed of

+ PMN and CFC areas (Sharma, 2011 for USD valuation)
+ Exploration & reserved areas of deep-sea mining (BUT these should be locked out rather than given a cost value)
+ Fishing effort (monetary value/index? using ready-made cost layer in Spatial Planning repo?)
+ Shipping routes (lock out? or use value of cell in Halpern et al., 2013 dataset?)

Need to define how to combine these costs into 1 layer. Mult-C Sync software (Moffett et al., 2005  / Cameron et al., 2008)?

Or: running multiple prioritization problems and defining **no-regret** MPAs across stakeholders.

*Update* possibly able to have USD value for fishing & PMN (see PMN subset)

### CFC

Cobalt-rich Ferromanganese Crusts. Still researching how to value.

```{r CFC, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
CFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp" 
PUxCFC <- fSpatPlan_Get_Polyg(CFC, PUs)  # function returns out as a logical vector
PUs <- PUs %>% 
  mutate(CFC = PUxCFC) # Add logical vector to variable called Mine1)Polyg1
```

### PMN

Areas rich in PolyMetallic Nodules. **Monetary value** could be assigned using:

+ density: 5.6kg/sqm in Indian Ocean basin (Sharma, 2011) - an approximation but better than nothing?
[link](https://www.resolve.ngo/docs/mar_technol_soc_j_45_28a.pdf)
+ value: 320-1,100USD/tonne of nodules (CRU Consulting, 2019) [link](https://isa.org.jm/files/files/documents/CRU_ISA%20Polymetallic%20nodule%20valuation%20report_24Aug2020.pdf)
+ size of PMN area: #of PUs that intersect with PMN layer * PU size (1000sqkm)
+ extraction rate: 1.5Mt/yr for 20 yrs (Sharma, 2011) (but using this brings even greater uncertainty, I think)

```{r PMN, echo = TRUE,  results = "hide", message = FALSE, warning = FALSE}
PMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp"
PUxPMN <- fSpatPlan_Get_Polyg(PMN, PUs)  
PUs <- PUs %>% 
  mutate(PMN = PUxPMN)
```

### Exploration areas

We need to **lock out** those areas, as mining companies have invested heavily into exploration and will not agree to lose access.

```{r expl, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
ExplPMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/01_pmn_exploration_areas.shp"
PUxExplPMN <- fSpatPlan_Get_Polyg(ExplPMN, PUs)  
PUs <- PUs %>% 
  mutate(ExplPMN = PUxExplPMN)

ExplPMS <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/02_pms_exploration_areas.shp"
PUxExplPMS <- fSpatPlan_Get_Polyg(ExplPMS, PUs)  
PUs <- PUs %>% 
  mutate(ExplPMS = PUxExplPMS)

ExplCFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/03_cfc_exploration_areas.shp"
PUxExplCFC <- fSpatPlan_Get_Polyg(ExplCFC, PUs)  
PUs <- PUs %>% 
  mutate(ExplCFC = PUxExplCFC)
```

### Reserved areas

Areas reserved for mining exploitation by developing countries - can also be locked out (same reasoning as above).

```{r res, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
ResPMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/04_pmn_reserved_areas.shp"
PUxResPMN <- fSpatPlan_Get_Polyg(ResPMN, PUs)  
PUs <- PUs %>% 
  mutate(ResPMN = PUxResPMN)

ResCFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/05_cfc_reserved_areas.shp"
PUxResCFC <- fSpatPlan_Get_Polyg(ResCFC, PUs)  
PUs <- PUs %>% 
  mutate(ResCFC = PUxResCFC)
```

Mining layer

*Still unable to add legend so showing the code, will ask on Slack*

```{r Mining layer, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
world <- ne_countries(scale = "medium", returnclass = "sf")
IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)
CFC <- st_read("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp")
PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp")
Expl_PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/01_pmn_exploration_areas.shp")
Expl_PMS <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/02_pms_exploration_areas.shp")
Expl_CFC <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/03_cfc_exploration_areas.shp")
Res_PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/04_pmn_reserved_areas.shp")
Res_CFC <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/05_cfc_reserved_areas.shp")
```

```{r Mining layer plot, echo = TRUE, message = FALSE, warning = FALSE}
ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = CFC, size = 1, color = "orange", fill = "orange") + 
  geom_sf(data = PMN, size = 1, color = "brown", fill = "brown") + 
  geom_sf(data = Expl_PMN, size = 1, color = "red", fill = "red") + 
  geom_sf(data = Expl_PMS, size = 1, color = "red", fill = "red") +
  geom_sf(data = Expl_CFC, size = 1, color = "red", fill = "red") +
  geom_sf(data = Res_PMN, size = 1, color = "red", fill = "red") +
  geom_sf(data = Res_CFC, size = 1, color = "red", fill = "red") +
  ggtitle("PMN, CFC, Exploration & Reserved areas") + 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)

```

## Shipping

**! cannot convert to polygon data when using code below (not written as code to avoid issues when knitting doc), error message says file too heavy** *will ask on Slack*

```

ShippingCost <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Shipping/shipping.tif") %>% 
    terra::as.polygons(trunc = FALSE, dissolve = FALSE, na.rm=FALSE) %>% # Convert to polygon data
    st_as_sf() %>% # Convert to sf
    st_transform(cCRS) %>% # Transform to projection
    st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
    rename(ShippingCost = layer)
```

```{r Shipping plot, echo = TRUE}
Shipping <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Shipping/shipping.tif")
plot(Shipping, main = 'Shipping routes')
```

## Fishing

```

Fishing <- fSpatPlan_Get_Cost(PUs, cCRS) ## if issues, try fx below

FishingCost <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Data/Cost/Cost_Raster_Sum.grd") %>% 
    terra::as.polygons(trunc = FALSE, dissolve = FALSE, na.rm=FALSE) %>% # Convert to polygon data
    st_as_sf() %>% # Convert to sf
    st_transform(cCRS) %>% # Transform to robinson
    st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
    rename(FishingCost = layer)

```
*Issues with plotting, should be a raster with colours based on cell USD value?*

```{r Fishing plot, echo = TRUE, message = FALSE, warnings = FALSE}
FishCost <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Data/Cost/Cost_Raster_Sum.grd")

plot(FishCost, main = 'Fishing cost')
```

# Prioritizr problems

## Basic run (Cons features)

*Assumes targets (ConsFeatures$amount) are set, code based on Alvise's work, not as code here to avoid issues when knitting*

```
p_ConsFeat <- problem(PUs, features = ConsFeatures, cost_column = "NoCost") %>% 
  add_min_set_objective() %>%
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%

````

### Cons features, fishing

*Assumes targets (ConsFeatures$amount) are set, code based on Alvise's work, not as code here to avoid issues when knitting*

```
p_ConsFeat_Fishing <- problem(PUs, features = ConsFeatures, cost_column = "Fishing") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
```

### Cons features, mining

*Assumes targets (ConsFeatures$amount) are set, code based on Alvise's work, not as code here to avoid issues when knitting*

*PMNCost will probably be created based on Sharma (2011) & CRU Consulting (2019)?*

```

p_ConsFeat_Mining <- problem(PUs, features = ConsFeatures, cost_column = "PMNCost") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
  add_locked_out_constraints("Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
```

### Cons features, mining + fishing

*Assumes targets (ConsFeatures$amount) are set, code based on Alvise's work, not as code here to avoid issues when knitting*

```

p_ConsFeat_MiningFishing <- problem(PUs, features = ConsFeatures, cost_column = "MiningFishing") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
  add_locked_out_constraints("Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
```

### Cons features, shipping

*Same as p_ConsFeat_MiningFishing but with shipping routes locked out as well*

*Not sure what kind of data transformation is needed to have a file with present/absent shipping routes*

*Or we can try estimating cost of shipping (question at start of .rmd) and combining with other costs into 1 layer*

```

p_ConsFeat_Shipping <- problem(PUs, features = ConsFeatures, cost_column = "MiningFishing") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
  add_locked_out_constraints("Shipping","Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
```

# Discussion

## Potential improvements

# Conclusion

### 2022-02-15 Msc_Thesis_Lea_v2.Rmd
Update Msc_Thesis_Lea_v2.Rmd

@@
 
+---
+
 # Table of Contents
@@
 + how to group all files (ex: all EBSAs + IBAs + IMMAs + vents + seamounts...) into one, to create ConsFeatures?
-+ should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection
-+ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what coupd be other methods (not using as.polygon)?
-+ shipping raster : cells have certain values (# of ships recorded??I will check for Tuesday), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)?
+Or is it enough to have columns next to PUs?
+RESOLVED: yes, stacking = creating columns next to PUs.
++ should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection. RESOLVED, yes, can use for this project (need to cross-check if want to publish)
++ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)? RESAMPLE if resolution to fine
++ shipping raster : cells have certain values (# of ships recorded??I can't figure from the attached files, metadata is confusing), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)?
 So we wouldn't try to compute the cost of deviating from any given route, but rather compute the price for each ship to go through 1 PU * the number of ships going through that PU.
 It sounds fairly logical to me, but maybe it's not?
-+ fishing cost layer : there is a ready-made fx in SpatPlan repo (thanks Jase & Tin), should we use it for this project? see fishing subsection (but issues with plotting  as you can see)
+RESOLVED: check 2008 metada, run prio problem with only shipping
++ fishing cost layer : there is a ready-made fx in SpatPlan repo (thanks Jase & Tin), should we use it for this project?
+see fishing subsection (but issues with plotting  as you can see).
+RESOLVED: yes, use fx from Jase & Tin.
+Change scale to remove outliers/crop to planning region.
++ I'm using echo = FALSE, message = FALSE, warning = FALSE, results = "hide" and still can't get rid of some coding output in the .html :( 
++ I st_to_sf the vents from .csv with PUs' crs but have some issues intersecting the points with polygs, will look more into it. ASK ON SLACK
+
+---
 
@@
 + to reach the global target of 30% marine protection, ABNJs are optimal (largest area; common good: higher incentive to protect since lower opportunity cost for individual countries)
++ but maybe we can also include non ABNJ? (because shipping data more interesting in EEZs, tbd later)
 
@@
 
-```{r, cache = FALSE, echo = FALSE, results = "hide", message = FALSE}
-#Open all the packages and get extras
+```{r, echo = FALSE, results = "hide", message = FALSE}
 options(repos = list(CRAN="http://cran.rstudio.com/")) 
@@
 
-still need to
++ **using the fishing cost layer  already created in Spatial Planning repo**
++ deep-sea features = seamounts, plateaus, vents
 
-+ **select specific fishing rasters (bycatch/demersal...)/ use the fishing cost layer  already created in Spatial Planning repo**
-+ and **select specific benthic shps (seamounts, ridges, plateaus, vents?)**
-+ **need to create subset without inactive vents** *filter fx not working because object=character, will ask on Slack*
+still need to intersect vent points with PUs
 
@@
  COST        |  Shipping              |  Halpern et al., 2013 (tif)
- COST        |  Fishing               |  Halpern et al., 2013 (tif)
+ COST        |  Fishing               |  Jase & Tin
  CONS        |  IBAs                  |  Birdlife datazone (shp)
@@
 
-Boundary
-```{r PR, echo=FALSE}
-ggplot(data = Bndry) + geom_sf()
-```
-
 Planning Units
@@
 + EBSAs
-+ Deepsea features: vents, seamounts, ridges, plateaus? *(need to define)*
++ Deepsea features: active vents, seamounts, plateaus?
 
@@
 
-**Still unable to add legend in plots so showing the code instead, but will ask on Slack**
+Seamounts,plateaus, vents
 
-Seamounts, ridges, plateaus, vents
-
-+ *need to add chunck for ridges & plateaus if we decide to add them*
-
-+ *need to check how to 'filter' out inactive vents and to cross vents (point data) with PUs and how to add in plot*
+*need to cross vents with PUs*
 
@@
 
-Ridges/Plateaus: >30%
+Plateaus: >30%
 
@@
 
+Plateaus <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/BlueHabs/Plateaus.shp"
+PUxPlateaus <- fSpatPlan_Get_Polyg(Plateaus, PUs)  
+PUs <- PUs %>% 
+  mutate(Plateaus = PUxPlateaus)
+
 Vents <- read.csv("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/vents.csv")
 VentsActive <- filter(Vents, Activity == "active, confirmed" | Activity == "active, inferred")
-PUxVents <- fSpatPlan_Get_Polyg(VentsActive, PUs)
-
+VentsActiveSF<- st_as_sf(VentsActive, coords = c("Latitude", "Longitude"), crs = st_crs(PUs))
+PUxVents <- st_intersects(PUs, VentsActiveSF) %>% 
+    as.logical() 
+PUs <- PUs %>% 
+  mutate(Vents = PUxVents) ##issue: NAs everywhere, probably because trying to cross points with polygs?
 ```
@@
 Seamounts <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/BlueHabs/Seamounts.shp")
+Plateaus <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/BlueHabs/Plateaus.shp")
 ```
 
-```{r Conservation features plot, echo = TRUE, message = FALSE, warning = FALSE}
+Conservation features
+
+```{r Conservation features plot, echo = FALSE, message = FALSE, warning = FALSE}
 
@@
   geom_sf(data = world) +
-  geom_sf(data = IBA, size = 1, color = "green", fill = "green") + 
-  geom_sf(data = IMMA, size = 1, color = "blue", fill = "blue")+ 
-  geom_sf(data = NEIO_09, size = 1, color = "grey", fill = "grey")+ 
-  geom_sf(data = NWIO_14, size = 1, color = "grey", fill = "grey")+
-  geom_sf(data = SIO_11, size = 1, color = "grey", fill = "grey")+
-  geom_sf(data = SIO_19, size = 1, color = "grey", fill = "grey")+
-  geom_sf(data = SIO_22, size = 1, color = "grey", fill = "grey")+
-  geom_sf(data = SIO_23, size = 1, color = "grey", fill = "grey")+
-  geom_sf(data = SIO_30, size = 1, color = "grey", fill = "grey")+
-  geom_sf(data = SIO_32, size = 1, color = "grey", fill = "grey")+
-  geom_sf(data = SIO_35, size = 1, color = "grey", fill = "grey")+
-  geom_sf(data = SIO_36, size = 1, color = "grey", fill = "grey")+
-  geom_sf(data = SIO_37, size = 1, color = "grey", fill = "grey")+
-  geom_sf(data = Seamounts, size = 1, color = "black", fill = "black")+
-  ggtitle("IBA, IMMA, EBSAs, Seamounts") + 
-  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
+  geom_sf(data = NEIO_09,  aes(color="A"), show.legend = "line")+ 
+  geom_sf(data = NWIO_14,  aes(color="A"), show.legend = "line")+
+  geom_sf(data = SIO_11,  aes(color="A"), show.legend = "line")+
+  geom_sf(data = SIO_19,  aes(color="A"), show.legend = "line")+
+  geom_sf(data = SIO_22,  aes(color="A"), show.legend = "line")+
+  geom_sf(data = SIO_23,  aes(color="A"), show.legend = "line")+
+  geom_sf(data = SIO_30,  aes(color="A"), show.legend = "line")+
+  geom_sf(data = SIO_32,  aes(color="A"), show.legend = "line")+
+  geom_sf(data = SIO_35,  aes(color="A"), show.legend = "line")+
+  geom_sf(data = SIO_36, aes(color="A"), show.legend = "line")+
+  geom_sf(data = SIO_37,  aes(color="A"), show.legend = "line")+
+  geom_sf(data = Seamounts,  aes(color="B"), show.legend = "line")+
+  geom_sf(data = Plateaus,  aes(color="C"), show.legend = "line")+
+  geom_sf(data = IBA,  aes(color="D"), show.legend = "line") + 
+  geom_sf(data = IMMA,  aes(color="E"), show.legend = "line")+ 
+  geom_sf(data = VentsActiveSF,  aes(color="F"), show.legend = "line")+ 
+  ggtitle("IBA, IMMA, EBSAs, Seamounts, Plateaus, Vents")+ 
+  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE) +
+  scale_color_manual(values = c("A" = "grey", "B" = "black", "C" = "purple", "D" = "green", "E" = "blue", "F" = "red"), 
+                     labels = c("EBSAs", "Seamounts", "Plateaus", "IBAs", "IMMAs", "Active vents"),
+                     name = "Legend")
+
 ```
-
 ## Cost layer
@@
 
-*Still unable to add legend so showing the code, will ask on Slack*
-
 ```{r Mining layer, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
@@
 
-```{r Mining layer plot, echo = TRUE, message = FALSE, warning = FALSE}
+Mining areas
+
+```{r Mining layer plot, echo = FALSE, message = FALSE, warning = FALSE}
 ggplot() + 
   geom_sf(data = world) +
-  geom_sf(data = CFC, size = 1, color = "orange", fill = "orange") + 
-  geom_sf(data = PMN, size = 1, color = "brown", fill = "brown") + 
-  geom_sf(data = Expl_PMN, size = 1, color = "red", fill = "red") + 
-  geom_sf(data = Expl_PMS, size = 1, color = "red", fill = "red") +
-  geom_sf(data = Expl_CFC, size = 1, color = "red", fill = "red") +
-  geom_sf(data = Res_PMN, size = 1, color = "red", fill = "red") +
-  geom_sf(data = Res_CFC, size = 1, color = "red", fill = "red") +
+  geom_sf(data = CFC, aes(color="A"), show.legend = "line") + 
+  geom_sf(data = PMN, aes(color="B"), show.legend = "line") + 
+  geom_sf(data = Expl_PMN, aes(color="C"), show.legend = "line") + 
+  geom_sf(data = Expl_PMS, aes(color="C"), show.legend = "line") +
+  geom_sf(data = Expl_CFC, aes(color="C"), show.legend = "line") +
+  geom_sf(data = Res_PMN, aes(color="D"), show.legend = "line") +
+  geom_sf(data = Res_CFC, aes(color="D"), show.legend = "line") +
+  scale_color_manual(values = c("A" = "orange", "B" = "brown", "C" = "red", "D" = "red"), 
+                     labels = c("CFC", "PMN", "Exploration areas", "Reserved areas"),
+                     name = "Legend") +
   ggtitle("PMN, CFC, Exploration & Reserved areas") + 
@@
 ```
-
 ## Shipping

### 2022-02-16 Msc_Thesis_Lea_v2.Rmd
Update Msc_Thesis_Lea_v2.Rmd
Added plots & histograms for fishing cost

@@
 ```{r, echo = FALSE, results = "hide", message = FALSE}
+#Open all the packages and get extras
 options(repos = list(CRAN="http://cran.rstudio.com/")) 
@@
  COST        |  Shipping              |  Halpern et al., 2013 (tif)
- COST        |  Fishing               |  Jase & Tin
+ COST        |  Fishing               |  Cost layer created by Jase
  CONS        |  IBAs                  |  Birdlife datazone (shp)
@@
 + EBSAs
-+ Deepsea features: active vents, seamounts, plateaus?
++ Deepsea features: active vents, seamounts, rift_valleys, plateaus? *(need to define)*
 
@@
 
-Conservation targets to be defined *need to write chunck to get ConsFeatures$amount*
+Conservation targets to be defined *need to write chunck*
 
@@
 
-Targets: *need to write chunck to get ConsFeatures$amount*
+Targets: *need to write chunck*
 
@@
 
-Seamounts,plateaus, vents
-
-*need to cross vents with PUs*
+Seamounts, plateaus, active vents
+Need to find out how to cross vents with PUs
 
@@
 
-Vents: >90%
+Vents: >90% [.csv to .shp](https://datacarpentry.org/r-raster-vector-geospatial/10-vector-csv-to-shapefile-in-r/)
 
-Seamounts: >70%
+Seamounts: >70% (ex: Letessier et al., 2019, PLOS Biology)
 
-Plateaus: >30%
+Ridges/Plateaus: >30%
 
@@
 VentsActive <- filter(Vents, Activity == "active, confirmed" | Activity == "active, inferred")
-VentsActiveSF<- st_as_sf(VentsActive, coords = c("Latitude", "Longitude"), crs = st_crs(PUs))
+VentsActiveSF<- st_as_sf(VentsActive, coords = c("Latitude", "Longitude"), crs = st_crs(PUs)) # could rewrite section using pipes
+```
+
+```
 PUxVents <- st_intersects(PUs, VentsActiveSF) %>% 
@@
 
-```{r Conservation features, echo = FALSE, message = FALSE, warning = FALSE}
+```{r Conservation features, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
 world <- ne_countries(scale = "medium", returnclass = "sf")
@@
 ```
+
 ## Cost layer
@@
 + PMN and CFC areas (Sharma, 2011 for USD valuation)
-+ Exploration & reserved areas of deep-sea mining (BUT these should be locked out rather than given a cost value)
-+ Fishing effort (monetary value/index? using ready-made cost layer in Spatial Planning repo?)
-+ Shipping routes (lock out? or use value of cell in Halpern et al., 2013 dataset?)
++ Exploration & reserved areas of deep-sea mining (lock out because won't agree to lose investments)
++ Fishing effort (using ready-made cost layer in Spatial Planning repo)
++ Shipping routes (use value of cell in Halpern et al., 2013 dataset for relative cost)
 
@@
 Or: running multiple prioritization problems and defining **no-regret** MPAs across stakeholders.
-
-*Update* possibly able to have USD value for fishing & PMN (see PMN subset)
 
@@
 PUs <- PUs %>% 
-  mutate(CFC = PUxCFC) # Add logical vector to variable called Mine1)Polyg1
+  mutate(CFC = PUxCFC) # Add logical vector to variable
 ```
@@
 
-Mining layer
-
 ```{r Mining layer, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
@@
   scale_color_manual(values = c("A" = "orange", "B" = "brown", "C" = "red", "D" = "red"), 
-                     labels = c("CFC", "PMN", "Exploration areas", "Reserved areas"),
+                     labels = c("CFC", "PMN", "Expl_PMN", "Expl_CFC", "Res_PMN", "Res_CFC"),
                      name = "Legend") +
@@
 ```
+
 ## Shipping
 
-**! cannot convert to polygon data when using code below (not written as code to avoid issues when knitting doc), error message says file too heavy** *will ask on Slack*
+**! cannot convert to polygon data when using code below (not written as code to avoid issues when knitting doc), error message says file too heavy** *Need to resample, will ask on Slack*
+
+```{r Shipping plot, echo = FALSE}
+Shipping <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Shipping/shipping.tif")
+plot(Shipping, main = 'Shipping routes')
+```
 
@@
 
-```{r Shipping plot, echo = TRUE}
-Shipping <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Shipping/shipping.tif")
-plot(Shipping, main = 'Shipping routes')
-```
-
 ## Fishing
 
-```
+Could set cut-off for cost outliers at 15,000 USD or 40,000 USD - depending on whether we want to favour variation or follow standard procedures for cut-off (see histograms and maps).
 
-Fishing <- fSpatPlan_Get_Cost(PUs, cCRS) ## if issues, try fx below
+```{r Fishing cost plot, echo = TRUE, message = FALSE, warnings = FALSE}
 
-FishingCost <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Data/Cost/Cost_Raster_Sum.grd") %>% 
+FishingCost <- fSpatPlan_Get_Cost <- function(PUs, cCRS){
+
+  # TODO Add some filtering ability to get the cost only for specific groups
+
+  Cost <- terra::rast("Data/Cost/Cost_Raster_Sum.grd") %>% 
     terra::as.polygons(trunc = FALSE, dissolve = FALSE, na.rm=FALSE) %>% # Convert to polygon data
@@
     st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
-    rename(FishingCost = layer)
+    rename(Cost = layer)
+}
 
-```
-*Issues with plotting, should be a raster with colours based on cell USD value?*
+hist(FishingCost$Cost) #max value >200,000
 
-```{r Fishing plot, echo = TRUE, message = FALSE, warnings = FALSE}
-FishCost <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Data/Cost/Cost_Raster_Sum.grd")
+ggplot() + 
+  geom_sf(data = world) +
+  geom_sf(data = FishingCost, aes(color=Cost)) +
+ggtitle("Fishing Cost") + 
+  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
 
-plot(FishCost, main = 'Fishing cost')
+FishingCostOutliers <- filter(FishingCost, Cost >= 50000) # check out outliers, play with cut-off value
+
+hist(FishingCostOutliers$Cost)
+
+ggplot() + 
+  geom_sf(data = world) +
+  geom_sf(data = FishingCostOutliers, aes(color=Cost)) +
+ggtitle("Fishing Cost Outliers (>50,000 USD)") + 
+  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
+
+FishingCostNoOutliers <- filter(FishingCost, Cost <= 50000) # remove outliers, play with cut-off value
+
+hist(FishingCostNoOutliers$Cost) # 49256 obs. = 524 outliers removed when cut-off = 20,000) but would also make sense to set cut-off at 15,000 or 40,000...
+
+ggplot() + 
+  geom_sf(data = world) +
+  geom_sf(data = FishingCostNoOutliers, aes(color=Cost)) +
+ggtitle("Fishing Cost Without Outliers") + 
+  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
+
 ```
@@
 ```
-p_ConsFeat_Fishing <- problem(PUs, features = ConsFeatures, cost_column = "Fishing") %>% 
+p_ConsFeat_Fishing <- problem(PUs, features = ConsFeatures, cost_column = "FishingCost") %>% 
   add_min_set_objective() %>% 
@@
 *PMNCost will probably be created based on Sharma (2011) & CRU Consulting (2019)?*
+
+[prioritizr.net](https://prioritizr.net/reference/add_locked_out_constraints.html) says to use structure below to lock out polygons but need to look more into it
+
+```
+# create problem with added locked out constraints using spatial polygon data
+locked_out <- sim_pu_polygons[sim_pu_polygons$locked_out == 1, ]
+p5 <- p1 %>% add_locked_out_constraints(locked_out)
+```
 

### 2022-02-19 Msc_Thesis_Lea_v2.Rmd
Update Msc_Thesis_Lea_v2.Rmd
work on fishing cost (histograms, maps) and vents (filtered, st_transformed, plotted)

@@
 
-Current questions:
+Resolved questions:
 
@@
 + should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection. RESOLVED, yes, can use for this project (need to cross-check if want to publish)
-+ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)? RESAMPLE if resolution to fine
 + shipping raster : cells have certain values (# of ships recorded??I can't figure from the attached files, metadata is confusing), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)?
@@
 Change scale to remove outliers/crop to planning region.
-+ I'm using echo = FALSE, message = FALSE, warning = FALSE, results = "hide" and still can't get rid of some coding output in the .html :( 
-+ I st_to_sf the vents from .csv with PUs' crs but have some issues intersecting the points with polygs, will look more into it. ASK ON SLACK
++ I'm using echo = FALSE, message = FALSE, warning = FALSE, results = "hide" and still can't get rid of some coding output in the .html. RESOLVED: cut up chuncks of code with separate ones for plots
++ I st_to_sf the vents from .csv with PUs' crs but have some issues intersecting the points with polygs, will look more into it. RESOLVED: use st_contains after crs = 4326 & st_transform(cCRS)
+
+---
+
+Current questions:
+
++ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)? RESAMPLE if resolution <1°. NEED TO FIND RESOLUTION of raster
 
@@
  COST        |  PMN & CFC (mining)    |  Petersen et al., 2016 (shp)
- COST        |  Shipping              |  Halpern et al., 2013 (tif)
+ COST        |  Shipping              |  Halpern et al., 2015 (tif from 2013)
  COST        |  Fishing               |  Cost layer created by Jase
@@
 
-Vents: >90% [.csv to .shp](https://datacarpentry.org/r-raster-vector-geospatial/10-vector-csv-to-shapefile-in-r/)
+Vents: >90% [.csv to .shp](https://datacarpentry.org/r-raster-vector-geospatial/10-vector-csv-to-shapefile-in-r/) [dataset  datum](https://vents-data.interridge.org/sites/vents-data.interridge.org/files/InterRidge_Vents_Database_Version3.1_documentation.pdf)
 
@@
 VentsActive <- filter(Vents, Activity == "active, confirmed" | Activity == "active, inferred") 
-VentsActiveSF<- st_as_sf(VentsActive, coords = c("Latitude", "Longitude"), crs = st_crs(PUs)) # could rewrite section using pipes
+VentsActiveSF <- st_as_sf(VentsActive, coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>% 
+  st_transform(cCRS)
+
+PUxVents <- st_contains(PUs,VentsActiveSF, sparse = FALSE) %>%
+  rowSums() %>% 
+  as.logical()
+PUs <- PUs %>% 
+  mutate(Vents = PUxVents)
+
 ```
 
-```
-PUxVents <- st_intersects(PUs, VentsActiveSF) %>% 
-    as.logical() 
-PUs <- PUs %>% 
-  mutate(Vents = PUxVents) ##issue: NAs everywhere, probably because trying to cross points with polygs?
+```{r vents plot, echo = FALSE, message = FALSE, warning = FALSE}
+world <- ne_countries(scale = "medium", returnclass = "sf")
+IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)
+
+ggplot() + 
+  geom_sf(data = world) + 
+  geom_sf(data = VentsActiveSF) +
+  ggtitle("Active vents after st_tranform(cCRS)") +
+  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
+
+#PUxVents <- st_join(PUs, VentsActiveSF, join = st_within) #no longer needed
+#VentsActiveSFClean <- dplyr::select(VentsActiveSF, Name.ID, geometry)
+#VentsActiveSFCleanBuff <- st_buffer(VentsActiveSFClean, dist = 1000)
+#PUxVents <- st_intersects(VentsActiveSF, PUs)
+#PUs <- PUs %>% 
+#  mutate(Vents = PUxVents)
+
 ```
@@
 ```{r Conservation features, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
-world <- ne_countries(scale = "medium", returnclass = "sf")
-IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)
 
@@
   ggtitle("IBA, IMMA, EBSAs, Seamounts, Plateaus, Vents")+ 
-  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE) +
+  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)+
   scale_color_manual(values = c("A" = "grey", "B" = "black", "C" = "purple", "D" = "green", "E" = "blue", "F" = "red"), 
@@
 
-```{r Fishing cost plot, echo = TRUE, message = FALSE, warnings = FALSE}
+```{r Fishing cost, echo = TRUE, warning = FALSE}
 
-FishingCost <- fSpatPlan_Get_Cost <- function(PUs, cCRS){
+#adding modified fx here to ensure it uses the right path
+
+fSpatPlan_Get_Cost <- function(PUs, cCRS){
 
@@
 
-  Cost <- terra::rast("Data/Cost/Cost_Raster_Sum.grd") %>% 
+  FishingCost <- terra::rast("Data/Cost/Cost_Raster_Sum.grd") %>% 
     terra::as.polygons(trunc = FALSE, dissolve = FALSE, na.rm=FALSE) %>% # Convert to polygon data
@@
     st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
-    rename(Cost = layer)
+    rename(FishingCost = layer)
 }
 
-hist(FishingCost$Cost) #max value >200,000
+FishingCost <- fSpatPlan_Get_Cost(PUs, cCRS)
+
+hist(FishingCost$FishingCost) #max value >200,000
+FishingCostPlusOne <- FishingCost$FishingCost + 1
+hist(log(FishingCostPlusOne)) #drop after log=10, exp(10)=22026.465794807
 
@@
   geom_sf(data = world) +
-  geom_sf(data = FishingCost, aes(color=Cost)) +
-ggtitle("Fishing Cost") + 
+  geom_sf(data = FishingCost, aes(color=FishingCost)) +
+ggtitle("Fishing Cost All Values") + 
   coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
 
-FishingCostOutliers <- filter(FishingCost, Cost >= 50000) # check out outliers, play with cut-off value
+```
+
+```
+FishingCostOutliers <- filter(FishingCost, Cost >= 22000) # check outliers
 
@@
 
-FishingCostNoOutliers <- filter(FishingCost, Cost <= 50000) # remove outliers, play with cut-off value
+FishingCostNoOutliers <- filter(FishingCost, Cost <= 22000) # remove outliers
 
-hist(FishingCostNoOutliers$Cost) # 49256 obs. = 524 outliers removed when cut-off = 20,000) but would also make sense to set cut-off at 15,000 or 40,000...
+hist(FishingCostNoOutliers$Cost) #(49256 obs. = 524 outliers removed when #cut-off = 20,000)
 

### 2022-02-23 Msc_Thesis_Lea_v2.Rmd
Update Msc_Thesis_Lea_v2.Rmd
added research question ideas, recap section, inactive vents,  some shipping exactextractr but not conclusive yet

@@
 author: "Lea Fourchault"
-date: "02/02/2022"
+date: "22/02/2022"
 output: html_document
@@
 ### Deep-sea features : seamounts, ridges, plateaus, vents 
+### Targets dataframe
 ## Cost layer
@@
 + should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection. RESOLVED, yes, can use for this project (need to cross-check if want to publish)
-+ shipping raster : cells have certain values (# of ships recorded??I can't figure from the attached files, metadata is confusing), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)?
++ shipping raster : cells have certain values (# of ships recorded??
+I can't figure from the attached files, metadata is confusing), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)?
 So we wouldn't try to compute the cost of deviating from any given route, but rather compute the price for each ship to go through 1 PU * the number of ships going through that PU.
@@
 + I st_to_sf the vents from .csv with PUs' crs but have some issues intersecting the points with polygs, will look more into it. RESOLVED: use st_contains after crs = 4326 & st_transform(cCRS)
++ add list of species as cons_feats: EN, VU etc (Ant's email to mining colleague)? RESOLVED: fine-tune later
++ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)?
+RESAMPLE if resolution <1°.
+Resolution = 934m*934m (!
+1° at equator = 111km = 60 nm; [convert](https://www.johndcook.com/how_big_is_a_degree.html); resolution too fine so need resampling).
+RESOLVED: no resampling, overlay raster with PUs (exactextractr) to attribute mean raster cells value to PUs (no weighted mean because raster cells are all equal-area because Mollweide proj)
++ how to remove EEZs from PUs? RESOLVED: find EEZ shp and spatial predicates, inverse overlay with Bndry, recreate PUs
++ setting  targets. RESOLVED: 30% for rpz area, 70% for important/rare feature; create dataframe with 1 column: features, 1 column: targets (match order)
 
@@
 
-Current questions:
+Current questions & issues:
 
-+ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)? RESAMPLE if resolution <1°. NEED TO FIND RESOLUTION of raster
++ ongoing - pricing deep-sea mineral resources (I am looking for more papers especially regarding CFC)
++ check research questions
++ overlay shipping & PUs + plot with log10?
++ cross inactive vents & PUs - done
++ how to assign price to mining shps
++ assigning targets - write chunk
+
+---
+
+Parameters we agree on for first prioritizr run
+
++ PUs = ABNJ only
++ cons feats = IMMAs, IBAs, chosen EBSAs, vents (active + inactive), seamounts, plateaus; no aquamaps
++ mining cost = value/t * density * area, discount rate and other improvements to be determined later (if any), USD
++ fishing cost = removed fish mass * fish value (layer created by Jase), USD
++ shipping cost = mean of shipping raster cells' values for each PU, not USD (intensity)
++ targets = 30% for rpz area, 70% for important
 
@@
 
-This project aims to identify areas for MPAs that maximise the conservation of biologically and ecologically important areas in the Indian Ocean's ABNJS, across all depths, while minimizing conflict with local industrial stakeholders - including the fishing, shipping, and deep-sea mining industries.
+*This project aims to identify areas for MPAs that maximise the conservation of biologically and ecologically important areas in the Indian Ocean's ABNJS, across all depths, while minimizing conflict with local industrial stakeholders - including the fishing, shipping, and deep-sea mining industries.*
+
+**Possible research questions**
+
+Ranked from most general to most precise, can be same question with different wording (some are in passive form, I know it's better to use active but some sound weird when saying 'we'?)
+
+1. how can marine spatial planning adapt to a time of increased anthropogenic pressures?
+2. how can the interests of multiple industrial stakeholders be incorporated into MPA design?
+3. how can value be estimated in the context of data-poor industries?
+4. which areas are most valuable to industrial stakeholders in the Indian Ocean?
+5. how can we design MPAs that minimize trade-offs with industrial stakeholders in the Indian Ocean?
+6. how can we design MPAs that maximize biodiversity conservation across depth layers in the Indian Ocean?
+7. how can we quantify/estimate trade-offs amongst industries in the Indian Ocean?
+8. how can we quantify/estimate trade-offs between industries and conservation in the Indian Ocean? and,
+9. how can such trade-offs with industrial stakeholders be minimized when designing MPAs?
 
@@
 
-+ **using the fishing cost layer  already created in Spatial Planning repo**
-+ deep-sea features = seamounts, plateaus, vents
-
-still need to intersect vent points with PUs
-
 blue habitats **do not include hydro vents**, so will use Beaulieu & Szanfranski (2020) [link](https://doi.pangaea.de/10.1594/PANGAEA.917894?format=html#download)
@@
  CONS        |  Longhurst provinces ? |  Marine Regions (shp)
- CONS        |  Vents                 |  Beaulieu & Szanfranski (2020) (csv)
+ CONS        |  Active vents          |  Beaulieu & Szanfranski (2020) (csv)
+ CONS        |  Inactive vents        |  Beaulieu & Szanfranski (2020) (csv)
 
 ## Planning region
+
+*substract EEZs shp from boundary, then recreate PUs*
 
@@
 + EBSAs
-+ Deepsea features: active vents, seamounts, rift_valleys, plateaus? *(need to define)*
++ Deepsea features: active + inactive vents, seamounts, plateaus
 
@@
 
-Conservation targets to be defined *need to write chunck to get ConsFeatures$amount*
+Conservation targets = 30%
 
@@
 
-Conservation targets to be defined *need to write chunck*
+Conservation targets = 30%
 
@@
 
-+ NEIO_09 : 10-30%
-+ NWIO_14 : 10-30%
-+ SIO_11 : 10-30%
-+ SIO_19 : >70%
-+ SIO_22 : >70%
-+ SIO_23 : >70%
-+ SIO_30 : >70%
-+ SIO_32 : >70%
-+ SIO_35 : >70%
-+ SIO_36 : >70%
-+ SIO_37 : >70%
++ NEIO_09 : 30%
++ NWIO_14 : 30%
++ SIO_11 :  30%
++ SIO_19 :  70%
++ SIO_22 :  70%
++ SIO_23 :  70%
++ SIO_30 :  70%
++ SIO_32 :  70%
++ SIO_35 :  70%
++ SIO_36 :  70%
++ SIO_37 :  70%
 
@@
 
-Seamounts, plateaus, active vents
-Need to find out how to cross vents with PUs
+Seamounts, plateaus, inactive + active vents
 
-Targets: tbd
-
-Vents: >90% [.csv to .shp](https://datacarpentry.org/r-raster-vector-geospatial/10-vector-csv-to-shapefile-in-r/) [dataset  datum](https://vents-data.interridge.org/sites/vents-data.interridge.org/files/InterRidge_Vents_Database_Version3.1_documentation.pdf)
-
-Seamounts: >70% (ex: Letessier et al., 2019, PLOS Biology)
-
-Ridges/Plateaus: >30%
++ Active + inactive vents: 70% [.csv to .shp](https://datacarpentry.org/r-raster-vector-geospatial/10-vector-csv-to-shapefile-in-r/) [dataset  datum](https://vents-data.interridge.org/sites/vents-data.interridge.org/files/InterRidge_Vents_Database_Version3.1_documentation.pdf)
++ Seamounts: 30% (but some advocate for more: Letessier et al., 2019, PLOS Biology)
++ Plateaus: >30%
 
@@
 
-PUxVents <- st_contains(PUs,VentsActiveSF, sparse = FALSE) %>%
+PUxVentsActive <- st_contains(PUs,VentsActiveSF, sparse = FALSE) %>%
   rowSums() %>% 
@@
 PUs <- PUs %>% 
-  mutate(Vents = PUxVents)
+  mutate(Active_Vents = PUxVentsActive)
+
+VentsInactive <- filter(Vents, Activity == "inactive")
+VentsInactiveSF <- st_as_sf(VentsActive, coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>% 
+  st_transform(cCRS)
+
+PUxVentsInactive <- st_contains(PUs,VentsInactiveSF, sparse = FALSE) %>%
+  rowSums() %>% 
+  as.logical()
+PUs <- PUs %>% 
+  mutate(Inactive_Vents = PUxVentsInactive)
 
@@
   geom_sf(data = VentsActiveSF,  aes(color="F"), show.legend = "line")+
+  geom_sf(data = VentsInactiveSF,  aes(color="G"), show.legend = "line")+
   ggtitle("IBA, IMMA, EBSAs, Seamounts, Plateaus, Vents")+ 
   coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)+
-  scale_color_manual(values = c("A" = "grey", "B" = "black", "C" = "purple", "D" = "green", "E" = "blue", "F" = "red"), 
-                     labels = c("EBSAs", "Seamounts", "Plateaus", "IBAs", "IMMAs", "Active vents"),
+  scale_color_manual(values = c("A" = "grey", "B" = "black", "C" = "blue", "D" = "green", "E" = "purple", "F" = "red", "G" = "brown"), 
+                     labels = c("EBSAs", "Seamounts", "Plateaus", "IBAs", "IMMAs", "Active vents", "Inactive Vents"),
                      name = "Legend")
@@
 ```
+### Targets dataframe
 
@@
 + Fishing effort (using ready-made cost layer in Spatial Planning repo)
-+ Shipping routes (use value of cell in Halpern et al., 2013 dataset for relative cost)
++ Shipping routes (use value of cell in Halpern et al. (2015), 2013 dataset for relative cost)
 
@@
   geom_sf(data = Expl_PMN, aes(color="C"), show.legend = "line") + 
-  geom_sf(data = Expl_PMS, aes(color="C"), show.legend = "line") +
-  geom_sf(data = Expl_CFC, aes(color="C"), show.legend = "line") +
-  geom_sf(data = Res_PMN, aes(color="D"), show.legend = "line") +
-  geom_sf(data = Res_CFC, aes(color="D"), show.legend = "line") +
-  scale_color_manual(values = c("A" = "orange", "B" = "brown", "C" = "red", "D" = "red"), 
-                     labels = c("CFC", "PMN", "Expl_PMN", "Expl_CFC", "Res_PMN", "Res_CFC"),
+  geom_sf(data = Expl_PMS, aes(color="D"), show.legend = "line") +
+ # geom_sf(data = Expl_CFC, aes(color="E"), show.legend = "line") +
+  geom_sf(data = Res_PMN, aes(color="F"), show.legend = "line") +
+ # geom_sf(data = Res_CFC, aes(color="G"), show.legend = "line") +
+  scale_color_manual(values = c("A" = "grey", "B" = "brown", "C" = "orange", "D" = "red", "F" = "yellow"), 
+                     labels = c("CFC", "PMN", "Expl_PMN", "Expl_PMS", "Res_PMN"),
                      name = "Legend") +
@@
 
-**! cannot convert to polygon data when using code below (not written as code to avoid issues when knitting doc), error message says file too heavy** *Need to resample, will ask on Slack*
+Overlay shipping raster with polygons to get average intensity of shipping for 1 PU.
 
@@
 plot(Shipping, main = 'Shipping routes')
+
+#ShippingRobin <- terra::
+#exactextractr::exact_extract(Shipping, PUs, mean)
 ```
@@
 
-Could set cut-off for cost outliers at 15,000 USD or 40,000 USD - depending on whether we want to favour variation or follow standard procedures for cut-off (see histograms and maps).
+Could set cut-off for cost outliers at 15,000 USD or 40,000 USD or no cut off - depending on whether we want to favour variation or follow standard procedures for outliers (see histograms and maps).
 
@@
 FishingCostPlusOne <- FishingCost$FishingCost + 1
-hist(log(FishingCostPlusOne)) #drop after log=10, exp(10)=22026.465794807
+hist(log10(FishingCostPlusOne)) #drop after log10(x)=4, x=10^4=10,000?
 

### 2022-02-23 Msc_Thesis_Lea_v2.Rmd
Update Msc_Thesis_Lea_v2.Rmd
Added cons_feats tibble

@@
 + Seamounts: 30% (but some advocate for more: Letessier et al., 2019, PLOS Biology)
-+ Plateaus: >30%
++ Plateaus: 30%
 
@@
 ### Targets dataframe
+
+Cons_feats <- tibble(Features = c("IBA", "IMMA", "NEIO_09", "NWIO_14", "SIO_11", "SIO_19", "SIO_22", "SIO_23", "SIO_30", "SIO_32", "SIO_35", "SIO_36", "SIO_37","Seamounts", "Plateaus", "Active_Vents", "Inactive_Vents"), Targets = c(0.3, 0.3, 0.3, 0.3, 0.3, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.3, 0.3, 0.7, 0.7))
 
@@
 
-*Assumes targets (ConsFeatures$amount) are set, code based on Alvise's work, not as code here to avoid issues when knitting*
+*Code based on Alvise's work, not as code here to avoid issues when knitting*
 
 ```
-p_ConsFeat <- problem(PUs, features = ConsFeatures, cost_column = "NoCost") %>% 
+p_ConsFeats <- problem(PUs, Cons_feats$Features, cost_column = "NoCost") %>%
   add_min_set_objective() %>%
-  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
+  add_relative_targets(Cons_Feats$Targets) %>% #
   add_binary_decisions() %>%

### 2022-03-02 Msc_Thesis_Lea_v2.Rmd
R file because bugs in .rmd file
All code from .rmd + area calculation chunk for PMN and CFC in Indian Ocean

@@
 + add list of species as cons_feats: EN, VU etc (Ant's email to mining colleague)? RESOLVED: fine-tune later
-+ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)?
++ shipping raster (Halpern et al., 2015) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)?
 RESAMPLE if resolution <1°.
@@
 + setting  targets. RESOLVED: 30% for rpz area, 70% for important/rare feature; create dataframe with 1 column: features, 1 column: targets (match order)
++ cross inactive vents & PUs - done
 
@@
 + overlay shipping & PUs + plot with log10?
-+ cross inactive vents & PUs - done
-+ how to assign price to mining shps
-+ assigning targets - write chunk
++ how to assign price to mining shps / create cost layer?
++ assigning targets - Cons_feats tibble created
 
@@
 
+# define region and boundary
+
 cCRS <- "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs" # Robinson projection
@@
 
+# intersect bndry and high seas
+
+ABNJ <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/World_High_Seas_v1/High_Seas_v1.shp") 
+ABNJRobin <- st_transform(ABNJ, cCRS) # high seas in robin proj, does not work when using pipe
+ABNJRobBuff  <- st_buffer(ABNJRobin, 0) # need to create buffer of 0 to avoid Error in CPL_geos_op2(op, x, y) : Evaluation error: TopologyException: Input geom 0 is invalid: Self-intersection at or near point -> overwrite offending geometries
+BndryABNJ <- st_intersection(ABNJRobBuff, Bndry)
+
+#create PUs
+
 PU_size <- 1000 # in km2, but check resolution for this project?
@@
 MaxDepth <- 200 # but check for deep-sea?
-PUs <- fSpatPlan_Get_PlanningUnits(Bndry, world, PU_size, Shape)
+PUs <- fSpatPlan_Get_PlanningUnits(BndryABNJ, world, PU_size, Shape)
 
+### method with EEZ took forever and was not created the PUs I wanted, so I chose to keep high seas shp instead of using EEZ shp
+#EEZ <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Data/World_EEZ_v11/eez_v11.shp") 
+#EEZRobin <-  st_transform(EEZ, cCRS)
+#BndryABNJ <- st_difference(Bndry, EEZRobin)
+#PUs <- fSpatPlan_Get_PlanningUnits(Bndry, world, PU_size, Shape) # error : st_crs(x) == st_crs(y) is not TRUE ?
 ```
@@
 
-```{r vents plot, echo = FALSE, message = FALSE, warning = FALSE}
+```{r vents plot, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
 world <- ne_countries(scale = "medium", returnclass = "sf")
@@
 
+```{r Conservation features and targets, echo = FALSE, message = FALSE, warning = FALSE}
+
 Cons_feats <- tibble(Features = c("IBA", "IMMA", "NEIO_09", "NWIO_14", "SIO_11", "SIO_19", "SIO_22", "SIO_23", "SIO_30", "SIO_32", "SIO_35", "SIO_36", "SIO_37","Seamounts", "Plateaus", "Active_Vents", "Inactive_Vents"), Targets = c(0.3, 0.3, 0.3, 0.3, 0.3, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.3, 0.3, 0.7, 0.7))
+
+```
 
@@
 
-```{r Shipping plot, echo = FALSE}
+```{r Shipping plot, echo = TRUE}
+
 Shipping <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Shipping/shipping.tif")
-plot(Shipping, main = 'Shipping routes')
+#View(Shipping)
+#plot(Shipping, main = 'Shipping routes')
+ShippingRobin <- terra::project(Shipping, "ESRI:54030") #transform to robin proj
+#plot(ShippingRobin, main = 'Shipping routes')
+PUxShipping <- exactextractr::exact_extract(ShippingRobin, PUs, "mean") # attribute mean value of pixels in each PU to PU
+PUs <- PUs %>% 
+  mutate(ShippingIntensity = PUxShipping) # add column to PU data frame
 
-#ShippingRobin <- terra::
-#exactextractr::exact_extract(Shipping, PUs, mean)
+#ShippingPlusOne <- Shipping$shipping +1
+#ShippingLog <- log10(ShippingPlusOne)
+#hist(log10(ShippingPlusOne))  #Warning: [hist] a sample of0% of the cells was used (of which 21% was NA); meaning +1 didn't work?
+
+#ShippingRobin <- terra::project(Shipping, "cCRS") #Error: [project] cannot get output boundaries
+
+#ShippingRobin <- project(Shipping, "+proj=robin +lon_0=0 +x_0=0 +y_0=0 #+datum=WGS84 +units=m +no_defs") # no error but does not assign projection
+
+#ShippingRobin <- project(Shipping, "ESRI:54030", method = "near") #transform to robin proj # error about method "near"
+
 ```
@@
 ```
+# not using this method because file to heavy to use as.polygons
 
@@
 ## Fishing
-
-Could set cut-off for cost outliers at 15,000 USD or 40,000 USD or no cut off - depending on whether we want to favour variation or follow standard procedures for outliers (see histograms and maps).
 
@@
     st_as_sf() %>% # Convert to sf
-    st_transform(cCRS) %>% # Transform to robinson
+    st_transform(cCRS) %>% # transform to robinson
     st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
@@
 FishingCost <- fSpatPlan_Get_Cost(PUs, cCRS)
-
-hist(FishingCost$FishingCost) #max value >200,000
-FishingCostPlusOne <- FishingCost$FishingCost + 1
-hist(log10(FishingCostPlusOne)) #drop after log10(x)=4, x=10^4=10,000?
+PUs <- PUs %>% 
+  mutate(Fishing = FishingCost) #issues: no error but cannot check PU dataframe after this: Error in `[.data.frame`(x, i) : undefined columns selected; probably need to restart session
+hist(FishingCost$FishingCost) 
+#FishingCostPlusOne <- FishingCost$FishingCost + 1 #not needed, min >0 for PU=ABNJ
+hist(log10(FishingCost$FishingCost)) #drop after log10(x)=4, x=10^4=10,000? #no longer needed since in ABNJ max val = ##USD 5000, and most values are 10^1 - 10^2 USD
 
@@
 ```
+
+#section not needed anymore since no outliers when PUs=ABNJ
+
 FishingCostOutliers <- filter(FishingCost, Cost >= 22000) # check outliers
@@
 ```
-p_ConsFeats <- problem(PUs, Cons_feats$Features, cost_column = "NoCost") %>%
+
+p_ConsFeats <- problem(PUs, features = Cons_feats$Features) %>% #issues because no cost_column
   add_min_set_objective() %>%
-  add_relative_targets(Cons_Feats$Targets) %>% #
+  add_relative_targets(Cons_feats$Targets) %>% 
   add_binary_decisions() %>%
-  add_lpsymphony_solver(verbose = FALSE) #%>%
+  add_lpsymphony_solver(verbose = FALSE)
 
-````
+```
 
@@
 
-*Assumes targets (ConsFeatures$amount) are set, code based on Alvise's work, not as code here to avoid issues when knitting*
+*code based on Alvise's work, not as code here to avoid issues when knitting*
 
 ```
-p_ConsFeat_Fishing <- problem(PUs, features = ConsFeatures, cost_column = "FishingCost") %>% 
+
+p_ConsFeat_Fishing <- problem(PUs, features = Cons_feats$Features, cost_column = PUs$FishingCost) %>% 
   add_min_set_objective() %>% 
-  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
+  add_relative_targets(Cons_feats$Targets) %>% 
   add_binary_decisions() %>%
-  add_lpsymphony_solver(verbose = FALSE) #%>%
+  add_lpsymphony_solver(verbose = FALSE)
+
 ```
@@
 
-*Assumes targets (ConsFeatures$amount) are set, code based on Alvise's work, not as code here to avoid issues when knitting*
+*code based on Alvise's work, not as code here to avoid issues when knitting*
 
@@
 
+# create problem with added locked out constraints using spatial polygon data
+
 ```
-# create problem with added locked out constraints using spatial polygon data
+use:
+
 locked_out <- sim_pu_polygons[sim_pu_polygons$locked_out == 1, ]
@@
 ```
+or try
 
@@
 
-p_ConsFeat_Mining <- problem(PUs, features = ConsFeatures, cost_column = "PMNCost") %>% 
+p_ConsFeat_Mining <- problem(PUs, features = Cons_feats$Features, cost_column = PUs$PMNCost) %>% 
   add_min_set_objective() %>% 
-  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
+  add_relative_targets(Cons_feats$Targets) %>% 
   add_binary_decisions() %>%
-  add_lpsymphony_solver(verbose = FALSE) #%>%
+  add_lpsymphony_solver(verbose = FALSE) %>%
   add_locked_out_constraints("Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
@@
 
-p_ConsFeat_MiningFishing <- problem(PUs, features = ConsFeatures, cost_column = "MiningFishing") %>% 
+p_ConsFeat_MiningFishing <- problem(PUs, features = Cons_feats, cost_column = PUs$MiningFishing) %>% 
   add_min_set_objective() %>% 
-  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
+  add_relative_targets(Cons_features$Targets) %>%
   add_binary_decisions() %>%
-  add_lpsymphony_solver(verbose = FALSE) #%>%
+  add_lpsymphony_solver(verbose = FALSE) %>%
   add_locked_out_constraints("Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
@@
 
-*Same as p_ConsFeat_MiningFishing but with shipping routes locked out as well*
-
 *Not sure what kind of data transformation is needed to have a file with present/absent shipping routes*
@@
 
-```
+*here, Shipping = mean values of pixels overlaid on PUs, so intensity of shippings; not USD*
 
-p_ConsFeat_Shipping <- problem(PUs, features = ConsFeatures, cost_column = "MiningFishing") %>% 
+```{r Shipping problem, echo = TRUE, warning = FALSE}
+
+p_ConsFeat_Shipping <- problem(PUs, features = Cons_feats$Features, cost_column = "PUs$Shipping") %>% 
   add_min_set_objective() %>% 
-  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
+  add_relative_targets(Cons_feats$Targets) %>% 
   add_binary_decisions() %>%
   add_lpsymphony_solver(verbose = FALSE) #%>%
-  add_locked_out_constraints("Shipping","Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
+  #add_locked_out_constraints("Shipping","Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
 ```

### 2022-03-02 SpatPlan_Master_Lea.R

@@ -1,0 +1,308 @@
+#' ---
+#' title: "Create a Spatial Plan"
+#' author: "Jason D. Everett"
+#' affiliation: "UQ/CSIRO/UNSW"
+#' date: "Last compiled on `r format(Sys.time(), '%A %d %B %Y')`"
+#' output: 
+#'   html_document
+#' ---
+#'
+
+##+ setup, include=FALSE
+knitr::opts_chunk$set(warning=FALSE, cache=FALSE, message=FALSE)
+# knitr::opts_chunk$set(collapse = TRUE, comment = "", warning = "off")
+
+#' ## Overview
+#' This code has been written to simplify the process for running a _prioritizr_ analysis on a given region. It is still a work in progress so feel free to submit pull requests with new features and code improvements.
+#' 
+#' The code depends on `sf`, `terra`, `tidyverse`, `rnaturalearth`, `prioritizr`, `stars`, `patchwork`.    
+#' 
+#' To use this code, you will need to download and expand `MME1DATA-Q1215/SpatialPlanning/Data.zip` to the directory `GitHub/SpatialPlanning/Data/`. Note that the download is only 2GB, but the expanded data is 35 GB in size. If you need help subsetting the data to your region due to memory or HD space constraints, contact Jason.
+#' 
+#' ## Preliminaries 
+source("SpatPlan_Extras.R") # Load the extras, including functions and libraries
+# source("SpatPlan_Process_AquaMaps.R") # This script reprocesses AquaMaps. WARNING: Lots of time (10s hrs) and memory (10s GB)
+# source("SpatPlan_Process_MPAs.R") # Only run if you need to reprocess the MPA data. You will need the WDPA files
+
+if (!file.exists("Data")) {
+  stop("The Data folder does not exist at SpatialPlanning/Data. Please download from the RDM and then try again.")
+}
+
[…]

### 2022-03-03 Shp/Biodiv_features/IMMAs/iucn-imma.qmd
Data
Adding data folder except shipping

<!DOCTYPE qgis PUBLIC 'http://mrcc.com/qgis.dtd' 'SYSTEM'>
<qgis version="3.10.4-A Coruña">
  <identifier>iucn-imma.shp</identifier>
  <parentidentifier>
Important Marine Mammal Area (IMMA)</parentidentifier>
  <language>GBR</language>
  <type>dataset</type>
  <title>IUCN Joint SSC/WCPA Marine Mammal Protected Area
Task Force (IUCN MMPATF) Spatial Distribution of
Important Marine Mammal Areas (IMMA)</title>
  <abstract>The IUCN IMMA is a joint project between the IUCN Species Survival Commission (SSC) and World Commission on Protected Areas (WCPA). The IUCN IMMA layer indicates discrete portions of habitat, important to marine mammal species that have the potential to be delineated and managed for conservation. IMMAs consist of areas that may merit place-based protection and/or monitoring. IMMAs can be seen as a marine mammal layer indicative of biodiversity and potentially ecosystem health for consideration by governments, intergovernmental organisations, conservation groups, and the general public. 
The IUCN IMMA layer contains information, which describes indicates discrete portions of habitat, important to marine mammal species that have the potential to be delineated and managed for conservation. Information described include the name, location, marine mammal species of importance, the broad level and finer sub-divisions criteria used, and the area size of the IMMA.

Purpose of creation: The IUCN IMMA layer has been created to be compatible with and inform the identification of other area-based management tools and conservation prioritisation areas and include but are not restricted to Ecologically or Biologically Significant Area (EBSA), Key Biodiversity Area (KBA), Important Bird and Biodiversity Area (IBA), RAMSAR Convention Site (RAMSAR), Cetacean Critical Habitat (CCH), Biologically Important Area (BIA). IMMAs consist of areas that may merit place-based protection and/or monitoring and may be considered as a marine mammal layer indicative of biodiversity and potentially ecosystem health for consideration by governments, intergovernmental organisations, conservation groups, and the general public.

Creation methodology: This dataset is drawn from knowledge provided directly to the IUCN MMPATF IMMA Secretariat by experts but may also include some data consolidated from 3rd party sources where other information may not have been available. These areas have been assessed by regional experts and were further assessed by a panel of  independent reviewers. Where possible the IUCN MMPATF has  included a polygon for each IMMA sub-region, which has been independently assessed across separate regional expert identification workshops (dated 31.10.21). Guidance on the use of selection criteria for the identification of IMMAs is available for download via https://www.marinemammalhabitat.org/downloads/ </abstract>
  <keywords vocabulary="gmd:topicCategory">
    <keyword>Biota</keyword>
    <keyword>Environment</keyword>
    <keyword>Oceans</keyword>
  </keywords>
  <contact>
    <name>Dr. Michael J. Tetley</name>
    <organization>IUCN Joint SSC/WCPA Marine Mammal Protected Area Task Force
(IUCN MMPATF)</organization>
    <position>IMMA Data Coordinator</position>
    <voice>+44 1688 400 524</voice>
    <fax>NA</fax>
    <email>michael.tetley@imma-network.org</email>
    <role>distributor</role>
  </contact>
  <links>
    <link description="IMMA Website and eAtlas" url="https://www.marinemammalhabitat.org/imma-eatlas/" type="information" format="Website" mimeType="NA" name="IMMA eAtlas" size="11.1mb"/>
  </links>
  <history>Note that the map shown in the metadata sheet is using the October 2021 release. The IMMA e-Atlas shows the up to date and additional layers (e.g. areas of interest - AoI, candidate IMMAs - cIMMAs) not available for public download (https://www.marinemammalhabitat.org/imma-eatlas).

The IUCN IMMA dataset is not necessarily a complete representation of all the important areas that have been identified for a region; the quality of the IUCN IMMA depends on the accessibility of accurate, comprehensive, up-to-date expert information from data and knowledge holders.

Other regions not currently assessed within the IUCN IMMA will be addressed in future regional expert workshops. New information will be assessed on a decadal basis to determine new IMMAs in the region or re-examine existing IMMAs. Data are updated on a 6-12 monthly basis.
</history>
  <fees></fees>
  <constraints type="Use">IUCN MMPATF (2019) Global Dataset of Important Marine Mammal
Areas (IUCN IMMA). [Insert month/year of the version downloaded].
Made available under agreement on terms and conditions of use by
the IUCN Joint SSC/WCPA Marine Mammal Protected Areas Task
Force and accessible via the IMMA e-Atlas
http://www.marinemammalhabitat.org/imma-eatlas/</constraints>
  <constraints type="Use">IUCN MMPATF (2019) Global Dataset of Important Marine Mammal
Areas (IUCN IMMA). [Insert month/year of the version downloaded].
Made available under agreement on terms and conditions of use by
the IUCN Joint SSC/WCPA Marine Mammal Protected Areas Task
Force and accessible via the IMMA e-Atlas
http://www.marinemammalhabitat.org/imma-eatlas/</constraints>
  <rights>IUCN IMMA User Licence Agreement for the purposes of Scientific Research, Education, or Conservation. Please see the IUCN IMMA User Licence Agreement for full terms and conditions of use and
supply. Please contact the IMMA Data Coordinator for further conditions of commercial use michael.tetley@imma-network.org
</rights>
  <license>Attribution-NonCommercial
CC BY-NC</license>
  <encoding></encoding>
  <crs>
    <spatialrefsys>
      <wkt>GEOGCRS["WGS 84",DATUM["World Geodetic System 1984",ELLIPSOID["WGS 84",6378137,298.257223563,LENGTHUNIT["metre",1]]],PRIMEM["Greenwich",0,ANGLEUNIT["degree",0.0174532925199433]],CS[ellipsoidal,2],AXIS["geodetic latitude (Lat)",north,ORDER[1],ANGLEUNIT["degree",0.0174532925199433]],AXIS["geodetic longitude (Lon)",east,ORDER[2],ANGLEUNIT["degree",0.0174532925199433]],USAGE[SCOPE["unknown"],AREA["World"],BBOX[-90,-180,90,180]],ID["EPSG",4326]]</wkt>
      <proj4>+proj=longlat +datum=WGS84 +no_defs</proj4>
      <srsid>3452</srsid>
      <srid>4326</srid>
      <authid>EPSG:4326</authid>
      <description>WGS 84</description>
      <projectionacronym>longlat</projectionacronym>
      <ellipsoidacronym>EPSG:7030</ellipsoidacronym>
      <geographicflag>true</geographicflag>
    </spatialrefsys>
  </crs>
  <extent>
    <spatial maxx="180" maxz="0" dimensions="2" minz="0" crs="EPSG:4326" maxy="45.79455600000000004" miny="-78.68547588799999914" minx="-180"/>
    <temporal>
      <period>
        <start>2016-09-30T20:00:00Z</start>
        <end>2021-10-31T11:06:24Z</end>
      </period>
    </temporal>
  </extent>
</qgis>

### 2022-03-03 SpatPlan_Master_Lea_v2.R
Create SpatPlan_Master_Lea_v2.R

@@ -1,0 +1,498 @@
+#' ---
+#' title: "Create a Spatial Plan"
+#' author: "Jason D. Everett"
+#' affiliation: "UQ/CSIRO/UNSW"
+#' date: "Last compiled on `r format(Sys.time(), '%A %d %B %Y')`"
+#' output: 
+#'   html_document
+#' ---
+#'
+
+##+ setup, include=FALSE
+knitr::opts_chunk$set(warning=FALSE, cache=FALSE, message=FALSE)
+# knitr::opts_chunk$set(collapse = TRUE, comment = "", warning = "off")
+
+#' ## Overview
+#' This code has been written to simplify the process for running a _prioritizr_ analysis on a given region. It is still a work in progress so feel free to submit pull requests with new features and code improvements.
+#' 
+#' The code depends on `sf`, `terra`, `tidyverse`, `rnaturalearth`, `prioritizr`, `stars`, `patchwork`.    
+#' 
+#' To use this code, you will need to download and expand `MME1DATA-Q1215/SpatialPlanning/Data.zip` to the directory `GitHub/SpatialPlanning/Data/`. Note that the download is only 2GB, but the expanded data is 35 GB in size. If you need help subsetting the data to your region due to memory or HD space constraints, contact Jason.
+#' 
+#' Script below modified by Lea Fourchault for Cross-sectoral conservation of the Indian Ocean project
+#' 
+#' ## Preliminaries 
+source("SpatPlan_Extras.R") # Load the extras, including functions and libraries
+# source("SpatPlan_Process_AquaMaps.R") # This script reprocesses AquaMaps. WARNING: Lots of time (10s hrs) and memory (10s GB)
+# source("SpatPlan_Process_MPAs.R") # Only run if you need to reprocess the MPA data. You will need the WDPA files
+
+if (!file.exists("Data")) {
[…]

### 2022-03-09 Msc_Thesis_Lea_v2.Rmd
Update Msc_Thesis_Lea_v2.Rmd
added prioritizr problem solving and plotting

@@
 author: "Lea Fourchault"
-date: "22/02/2022"
-output: html_document
+date: "08/03/2022"
+output: 
+ html_document:
+    toc: true
+    toc_float: true
+
 ---
@@
 ```
-
----
-
-# Table of Contents
-# Introduction
-# Materials & Methods
-## Workflow with Prioritizr 
-## Overview of data  & layers
-## Planning region 
-## Conservation features & targets
-### IBAs
-### IMMAs
-### EBSAs
-### Deep-sea features : seamounts, ridges, plateaus, vents 
-### Targets dataframe
-## Cost layer
-### CFC
-### PMN
-### Exploration areas
-### Reserved areas
-### Shipping
-### Fishing *use cost layer already created?*
-# Prioritizr problems
-### Basic run (only Cons features)
-### Cons features, fishing
-### Cons features, mining (lock-out expl  & res areas; assign cost for PMN areas)
-### Cons features, fishing + mining 
-### Cons features, shipping 
-# Discussion
-## Potential improvements
-# Conclusion
-
 ---
@@
 + setting  targets. RESOLVED: 30% for rpz area, 70% for important/rare feature; create dataframe with 1 column: features, 1 column: targets (match order)
-+ cross inactive vents & PUs - done
++ *cross inactive vents & PUs - done*
++ check research questions. RESOLVED keep Ant's questions
++ Error in h(simpleError(msg, call)) : error in evaluating the argument 'x' in selecting a method for function 'add_relative_targets': error in evaluating the argument 'features' in selecting a method for function 'problem': object 'Cons_feats' not found.
+RESOLVED: use string: features = "x", "y"; instead of creating Cons_feats tibble
++ cannot install lpsymphony: almost compiled but failed because did not find directory for gfortran AND see this link: issues with Mojave https://www.cynkra.com/blog/2021-03-16-gfortran-macos/ : brew and gfortran cask deprecated since 2021 (still work if installed before but i didn't).
+RESOLVED: install Gurobi using student email
++ PUs: when mutating FishingCost column to PUs, creates duplicate of 'geometry' column called 'Fishing.geometry'.
+RESOLVED don't forget to add dataframe: dataframeDOLLARcolumn name: PUs <- PUs %>% mutate(Fishing = FishingCost$FishingCost)
++ *create new PUs with ABNJ only - done but check method*, methodology not checked
++ general bugs to run code on .rmd file.
+Use knitr::spin instead.
+I also replaced the initial chunk with ```{r setup, include=FALSE} knitr::opts_chunk$set(echo = TRUE)``` but no result, maybe need to reinstall r and r studio
 
@@
 
-Current questions & issues:
+Current questions & updates:
 
-+ ongoing - pricing deep-sea mineral resources (I am looking for more papers especially regarding CFC)
-+ check research questions
-+ overlay shipping & PUs + plot with log10?
-+ how to assign price to mining shps / create cost layer?
-+ assigning targets - Cons_feats tibble created
++ ongoing - pricing deep-sea mineral resources: for CFC *we know the area based on Crust.shp; we know CFC mean specific-surface areas = 325 m2/g (Hein et al., 2000; found in Hein et al., 2013) -> can find tonnage; then, we know chemical composition of CFC in Indian Ocean (given in wt% or ppm=g/tonne, for ex: 392 Mo ppm; Fe=22wt% - see Hein, 2014 or Petersen et al., 2016) -> can multiply by CFC tonnage to get tonnage for each mineral; then multiply by market prices (mean over 5 yrs?) to estimate total value of CFC in Indian Ocean?*
++ I have emailed Dr. Mizell about her calculations of CFC tonnage for the Indian Ocean (paper only gives global tonnage)
+
+*Coding questions, will think about it some more*
+
++ how to assign price to mining shps / create mining cost layer once value of deep-sea resources is computed?
++ overlay shipping & PUs + plot with log10: partly resolved using exact_extractr but issues with +1 and log10 and when plotting PUxShipping: not a map but a scatterplot; issues with transformation/reprojection?
 
@@
 + PUs = ABNJ only
-+ cons feats = IMMAs, IBAs, chosen EBSAs, vents (active + inactive), seamounts, plateaus; no aquamaps
-+ mining cost = value/t * density * area, discount rate and other improvements to be determined later (if any), USD
++ cons_feats = IMMAs, IBAs, chosen EBSAs, vents (active + inactive), seamounts, plateaus; no aquamaps
++ PMN mining value = value/t * density * area, discount rate and other improvements to be determined later (if any), USD
 + fishing cost = removed fish mass * fish value (layer created by Jase), USD
-+ shipping cost = mean of shipping raster cells' values for each PU, not USD (intensity)
-+ targets = 30% for rpz area, 70% for important
++ shipping cost = mean of shipping raster cells' values for each PU using exact_extracr, not USD (intensity) OR SHOULD WE USE SHIPPING AS A PENALTY?
++ targets = 30% for rpz area, 70% for important/rare feature
 
@@
 
+**Aim**
+
 *This project aims to identify areas for MPAs that maximise the conservation of biologically and ecologically important areas in the Indian Ocean's ABNJS, across all depths, while minimizing conflict with local industrial stakeholders - including the fishing, shipping, and deep-sea mining industries.*
 
-**Possible research questions**
+**Current research questions**
 
-Ranked from most general to most precise, can be same question with different wording (some are in passive form, I know it's better to use active but some sound weird when saying 'we'?)
-
-1. how can marine spatial planning adapt to a time of increased anthropogenic pressures?
-2. how can the interests of multiple industrial stakeholders be incorporated into MPA design?
-3. how can value be estimated in the context of data-poor industries?
-4. which areas are most valuable to industrial stakeholders in the Indian Ocean?
-5. how can we design MPAs that minimize trade-offs with industrial stakeholders in the Indian Ocean?
-6. how can we design MPAs that maximize biodiversity conservation across depth layers in the Indian Ocean?
-7. how can we quantify/estimate trade-offs amongst industries in the Indian Ocean?
-8. how can we quantify/estimate trade-offs between industries and conservation in the Indian Ocean? and,
-9. how can such trade-offs with industrial stakeholders be minimized when designing MPAs?
+1. How can future seabed mining be included in spatial planning?
+2. What is the potential spatial overlap among biodiversity, fishing, shipping and seabed mining in the Indian Ocean high seas?
+3. Where could we place reserves in the Indian Ocean high seas that meet conservation objectives whilst minimsing conflict with fishing and seabed mining?
 
@@
 
-blue habitats **do not include hydro vents**, so will use Beaulieu & Szanfranski (2020) [link](https://doi.pangaea.de/10.1594/PANGAEA.917894?format=html#download)
+blue habitats do not include hydro vents, so will use Beaulieu & Szanfranski (2020) [link](https://doi.pangaea.de/10.1594/PANGAEA.917894?format=html#download)
 
@@
   st_transform(cCRS)
-Region <- c(xmin = 18, xmax = 130, ymin = -39, ymax = 34) # Indian Ocean limits
+#Region <- c(xmin = 18, xmax = 140, ymin = -39, ymax = 34) 
+Region <- c(xmin = 20, xmax = 120, ymin = -60, ymax = 20) 
+# Indian Ocean limits: CIA World Factbook (but see: According to the International Hydrographic Organization (IHO) and the United Nations Oceans Atlas, the area from 40°S latitude to 60°S latitude is included in the Indian Ocean.
+The area that encircles the globe from 60°S latitude to the coast of Antarctica is called The Great Southern Ocean.
+The Indian Ocean's width extends from 45° E longitude to 110°East longitude.) Also, Region has big impact on final plan because fishing cost is much lower as latitude increases towards south pole
 Bndry <- fSpatPlan_Get_Boundary(Region, cCRS) # boundary
@@
 VentsInactive <- filter(Vents, Activity == "inactive")
-VentsInactiveSF <- st_as_sf(VentsActive, coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>% 
+VentsInactiveSF <- st_as_sf(VentsInactive, coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>% 
   st_transform(cCRS)
@@
 world <- ne_countries(scale = "medium", returnclass = "sf")
-IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)
+#IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)
+IndianO <- c(xmin = 20, xmax = 120, ymin = -60, ymax = 20)
 
-ggplot() + 
-  geom_sf(data = world) + 
-  geom_sf(data = VentsActiveSF) +
-  ggtitle("Active vents after st_tranform(cCRS)") +
-  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
+#ggplot() + 
+ # geom_sf(data = world) + 
+  #geom_sf(data = VentsInactiveSF) +
+  #ggtitle("Active vents after st_tranform(cCRS)") +
+  #coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
 
-#PUxVents <- st_join(PUs, VentsActiveSF, join = st_within) #no longer needed
+# no longer needed
+#PUxVents <- st_join(PUs, VentsActiveSF, join = st_within) 
 #VentsActiveSFClean <- dplyr::select(VentsActiveSF, Name.ID, geometry)
@@
 
+IndianO <- c(xmin = 20, xmax = 120, ymin = -60, ymax = 20)
+
 ggplot() + 
@@
   ggtitle("IBA, IMMA, EBSAs, Seamounts, Plateaus, Vents")+ 
-  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)+
+  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE)+
   scale_color_manual(values = c("A" = "grey", "B" = "black", "C" = "blue", "D" = "green", "E" = "purple", "F" = "red", "G" = "brown"), 
@@
 ```
-### Targets dataframe
 
-```{r Conservation features and targets, echo = FALSE, message = FALSE, warning = FALSE}
+### Longhurst provinces
 
-Cons_feats <- tibble(Features = c("IBA", "IMMA", "NEIO_09", "NWIO_14", "SIO_11", "SIO_19", "SIO_22", "SIO_23", "SIO_30", "SIO_32", "SIO_35", "SIO_36", "SIO_37","Seamounts", "Plateaus", "Active_Vents", "Inactive_Vents"), Targets = c(0.3, 0.3, 0.3, 0.3, 0.3, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.3, 0.3, 0.7, 0.7))
+```{r longhurst plot, echo = FALSE, message = FALSE, warning = FALSE}
+
+longh <- st_read(file.path("Data","LonghurstProvinces","Longhurst_world_v4_2010.shp")) %>% 
+  st_make_valid()
+longhRob <- st_transform(longh, cCRS)
+
+nr <- st_nearest_feature(PUs, longhRob)
+
+PUs <- PUs %>% 
+  mutate(ProvCode = longh$ProvCode[nr],
+         ProvDescr = longh$ProvDescr[nr])
+
+ggplot() + 
+  geom_sf(data = world, color = "grey20", fill = "grey20", size = 0.1, show.legend = FALSE) +
+  geom_sf(data = PUs, aes(fill = ProvCode), colour = "grey50", size = 0.05, alpha = 0.7, show.legend = TRUE) +
+  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE)
 
@@
 
-## Cost layer
+### Targets dataframe
 
-Composed of
-
-+ PMN and CFC areas (Sharma, 2011 for USD valuation)
-+ Exploration & reserved areas of deep-sea mining (lock out because won't agree to lose investments)
-+ Fishing effort (using ready-made cost layer in Spatial Planning repo)
-+ Shipping routes (use value of cell in Halpern et al. (2015), 2013 dataset for relative cost)
-
-Need to define how to combine these costs into 1 layer. Mult-C Sync software (Moffett et al., 2005  / Cameron et al., 2008)?
-
-Or: running multiple prioritization problems and defining **no-regret** MPAs across stakeholders.
-
-### CFC
-
-Cobalt-rich Ferromanganese Crusts. Still researching how to value.
-
-```{r CFC, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
-CFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp" 
-PUxCFC <- fSpatPlan_Get_Polyg(CFC, PUs)  # function returns out as a logical vector
-PUs <- PUs %>% 
-  mutate(CFC = PUxCFC) # Add logical vector to variable
 ```
 
-### PMN
+#Cons_feats <- tibble(Features = c("IBA", "IMMA", "NEIO_09", "NWIO_14", "SIO_11", "SIO_19", "SIO_22", "SIO_23", "SIO_30", "SIO_32", "SIO_35", "SIO_36", "SIO_37","Seamounts", "Plateaus", "Active_Vents", "Inactive_Vents"), Targets = c(0.3, 0.3, 0.3, 0.3, 0.3, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.3, 0.3, 0.7, 0.7))
 
-Areas rich in PolyMetallic Nodules. **Monetary value** could be assigned using:
-
-+ density: 5.6kg/sqm in Indian Ocean basin (Sharma, 2011) - an approximation but better than nothing?
-[link](https://www.resolve.ngo/docs/mar_technol_soc_j_45_28a.pdf)
-+ value: 320-1,100USD/tonne of nodules (CRU Consulting, 2019) [link](https://isa.org.jm/files/files/documents/CRU_ISA%20Polymetallic%20nodule%20valuation%20report_24Aug2020.pdf)
-+ size of PMN area: #of PUs that intersect with PMN layer * PU size (1000sqkm)
-+ extraction rate: 1.5Mt/yr for 20 yrs (Sharma, 2011) (but using this brings even greater uncertainty, I think)
-
-```{r PMN, echo = TRUE,  results = "hide", message = FALSE, warning = FALSE}
-PMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp"
-PUxPMN <- fSpatPlan_Get_Polyg(PMN, PUs)  
-PUs <- PUs %>% 
-  mutate(PMN = PUxPMN)
 ```
 
-### Exploration areas
+## Locked out areas
 
-We need to **lock out** those areas, as mining companies have invested heavily into exploration and will not agree to lose access.
+### polymetallic nodules / polymetallic sulfides: exploration & reserved areas
 
-```{r expl, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
+```{r Locked out, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
+
 ExplPMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/01_pmn_exploration_areas.shp"
@@
 
-ExplCFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/03_cfc_exploration_areas.shp"
-PUxExplCFC <- fSpatPlan_Get_Polyg(ExplCFC, PUs)  
-PUs <- PUs %>% 
-  mutate(ExplCFC = PUxExplCFC)
-```
-
-### Reserved areas
-
-Areas reserved for mining exploitation by developing countries - can also be locked out (same reasoning as above).
-
-```{r res, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
 ResPMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/04_pmn_reserved_areas.shp"
@@
 
-ResCFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/05_cfc_reserved_areas.shp"
-PUxResCFC <- fSpatPlan_Get_Polyg(ResCFC, PUs)  
-PUs <- PUs %>% 
-  mutate(ResCFC = PUxResCFC)
 ```
 
-```{r Mining layer, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
+```{r Plotting Locked out, echo = FALSE, message = FALSE, warning = FALSE}
+# plotting locked out areas
+
 world <- ne_countries(scale = "medium", returnclass = "sf")
-IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)
+IndianO <- c(xmin = 18, ymin = -60, xmax = 130, ymax = 22)
+Expl_PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/01_pmn_exploration_areas.shp")
+Expl_PMS <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/02_pms_exploration_areas.shp")
+Res_PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/04_pmn_reserved_areas.shp")
+
+ggplot() + 
+  geom_sf(data = world) +
+  geom_sf(data = Expl_PMN, aes(color="A"), show.legend = "line") + 
+  geom_sf(data = Expl_PMS, aes(color="B"), show.legend = "line") +
+  geom_sf(data = Res_PMN, aes(color="C"), show.legend = "line") +
+  scale_color_manual(values = c("A" = "red", "B" = "yellow", "C" = "black"),
+                     labels = c("Expl_PMN", "Expl_PMS", "Res_PMN"),
+                     name = "Legend") +
+  ggtitle("Exploration & Reserved areas") + 
+  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE)
+
+```
+
+## Cost layer
+
+Composed of
+
++ PMN and CFC areas (Sharma, 2011/ Mizell et al., 2022 for USD valuation) - in progress
++ Fishing effort (using ready-made cost layer in Spatial Planning repo)
++ Shipping routes (use value of cell in Halpern et al. (2015), 2013 dataset for relative cost) - in progress, but should be used as penalty, not cost?
+
+Need to define how to combine these costs into 1 layer. Mult-C Sync software (Moffett et al., 2005  / Cameron et al., 2008)?
+
+Or: running multiple prioritization problems and defining **no-regret** MPAs across stakeholders.
+
+### Mineral resources
+
+```{r Plot PMN CFC, echo = FALSE, message = FALSE, warning = FALSE}
+#' Plotting mining resources
+
+world <- ne_countries(scale = "medium", returnclass = "sf")
+#IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)
+IndianO <- c(xmin = 20, xmax = 120, ymin = -60, ymax = 20)
 CFC <- st_read("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp")
 PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp")
-Expl_PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/01_pmn_exploration_areas.shp")
-Expl_PMS <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/02_pms_exploration_areas.shp")
-Expl_CFC <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/03_cfc_exploration_areas.shp")
-Res_PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/04_pmn_reserved_areas.shp")
-Res_CFC <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/05_cfc_reserved_areas.shp")
-```
-
-Mining areas
-
-```{r Mining layer plot, echo = FALSE, message = FALSE, warning = FALSE}
 ggplot() + 
@@
   geom_sf(data = PMN, aes(color="B"), show.legend = "line") + 
-  geom_sf(data = Expl_PMN, aes(color="C"), show.legend = "line") + 
-  geom_sf(data = Expl_PMS, aes(color="D"), show.legend = "line") +
- # geom_sf(data = Expl_CFC, aes(color="E"), show.legend = "line") +
-  geom_sf(data = Res_PMN, aes(color="F"), show.legend = "line") +
- # geom_sf(data = Res_CFC, aes(color="G"), show.legend = "line") +
-  scale_color_manual(values = c("A" = "grey", "B" = "brown", "C" = "orange", "D" = "red", "F" = "yellow"), 
-                     labels = c("CFC", "PMN", "Expl_PMN", "Expl_PMS", "Res_PMN"),
+  scale_color_manual(values = c("A" = "grey", "B" = "brown"),
+                     labels = c("CFC", "PMN"),
                      name = "Legend") +
-  ggtitle("PMN, CFC, Exploration & Reserved areas") + 
-  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
+  ggtitle("Polymetallic Nodules and Cobalt-rich Ferromanganese Crust areas") + 
+  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE)
+
+```
+
+CFC value
+script to get area modified from https://rpubs.com/rural_gis/255550
+
+```{r CFC, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
+CFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp" 
+PUxCFC <- fSpatPlan_Get_Polyg(CFC, PUs)  # function returns out as a logical vector
+PUs <- PUs %>% 
+  mutate(CFC = PUxCFC)
+
+#load CFC areas polygon shapefile 
+CFC <- st_read("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp")
+CFCRobin <- st_transform(CFC, cCRS) #robin proj
+
+#run the intersect function, converting the output to a tibble in the process
+#int <- as_tibble(st_intersection(CFCRobin, BndryABNJ))
+int <- st_intersection(CFCRobin, BndryABNJ)
+
+#add in an area count column to the tibble (area of each CFC poly in intersect layer)
+int$CFCarea <- st_area(int)
+```
+
+Intersections between PUs and CFC in red - correct
+
+```{r plot CFC, echo = FALSE, message = FALSE, warning = FALSE}
+
+#plot the layers to visually check result of intersect
+plot (CFCRobin$geometry, col='green')
+plot(BndryABNJ$geometry, add=T)
+plot(int, col='red', add=T) #correct
+
+#group data by county area and calculate the total arable land area per county
+#output as new tibble
+AreaCFCtotal <- int %>%
+  summarise(areaCFCtot = sum(CFCarea))
+```
+
+```
+# CFC value code (so you can have a look)
+# script to get area modified from https://rpubs.com/rural_gis/255550
+
+CFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp" 
+PUxCFC <- fSpatPlan_Get_Polyg(CFC, PUs)  # function returns out as a logical vector
+PUs <- PUs %>% 
+  mutate(CFC = PUxCFC)
+
+#load CFC areas polygon shapefile 
+CFC <- st_read("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp")
+CFCRobin <- st_transform(CFC, cCRS) #robin proj
+
+#run the intersect function, converting the output to a tibble in the process
+#int <- as_tibble(st_intersection(CFCRobin, BndryABNJ))
+int <- st_intersection(CFCRobin, BndryABNJ)
+
+#add in an area count column to the tibble (area of each CFC poly in intersect layer)
+int$CFCarea <- st_area(int)
+
+#plot the layers to visually check result of intersect
+plot (CFCRobin$geometry, col='green')
+plot(BndryABNJ$geometry, add=T)
+plot(int, col='red', add=T) #correct
+
+#group data by county area and calculate the total arable land area per county
+#output as new tibble
+AreaCFCtotal <- int %>%
+  summarise(areaCFCtot = sum(CFCarea)) # = 2.256596e+12 [m^2]
+```
+**CFC area = 2.256596e+12 [m^2], seems a lot (see code above, no error messages but probably wrong somewhere)? Would be 2256596 sqkm, and I have 39803 PUs of 1000 sqkm each**
+
+*Other method to check :* 
+area of CFC in PUs < 2753*1000 sqkm = 2.753e+6 sqm (see sum of TRUE intersections between PUs and CFC poly below)
+
+```
+sum(PUxCFC) # sum of cross = TRUE between PUs and CFC polyg
+[1] 2753
+
+```
+PMN - same code
+
+```{r PMN, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
+
+# PMN value
+
+#script to get area modified from https://rpubs.com/rural_gis/255550
+
+#load county areas polygon shapefile
+
+PMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp"
+PUxPMN <- fSpatPlan_Get_Polyg(PMN, PUs)  
+PUs <- PUs %>% 
+  mutate(PMN = PUxPMN)
+
+PMN <- st_read("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp")
+PMNRobin <- st_transform(PMN, cCRS) #robin proj
+
+#run the intersect function, converting the output to a tibble in the process
+int2 <- st_intersection(PMNRobin, BndryABNJ) #need this to avoid Error in UseMethod("st_area") : 
+#int2 <- as_tibble(st_intersection(PMNRobin, BndryABNJ)) #no applicable method for 'st_area' applied to an object of class "c('tbl_df', 'tbl', 'data.frame')"
+
+#add in an area count column to the tibble (area of each CFC poly in intersect layer)
+int2$PMNarea <- st_area(int2)
+```
+
+Intersection between PUs and PMN in red - correct
+
+```{r plot PMN, echo = FALSE, message = FALSE, warning = FALSE}
+
+#plot the layers to visually check result of intersect
+plot (PMNRobin$geometry, col='green')
+plot(BndryABNJ$geometry, add=T)
+plot(int2, col='red', add=T) #correct
+
+#group data by county area and calculate the total arable land area per county
+#output as new tibble
+AreaPMNtotal <- int2 %>%
+  summarise(areaPMNtot = sum(PMNarea)) #1.933646e+12 [m^2],  seems a lot?
+```
+
+**Same issue for PMN as for CFC: intersection area too high: gives 1.933646e+12 [m^2],  seems a lot?**
+
+*other method to check*
+area of PMN in PUs < 2075*1000 sqkm = 2.075e+6 sqm (see sum of TRUE intersections between PUs and CFC poly below)
+
+```
+# area of PMN in PUs should be < 2075*1000 sqkm because:
+ sum(PUxPMN)
+ [1] 2075
+# and 1 PU = 1000 sqkm
+### Fishing cost
+
+```{r Fishing cost, echo = FALSE, message = FALSE, warning = FALSE}
+
+#Fishing cost
+
+FishingCost <- terra::rast("Data/Cost/Cost_Raster_Sum.grd") %>% 
+  terra::as.polygons(trunc = FALSE, dissolve = FALSE, na.rm=FALSE) %>% # Convert to polygon data
+  st_as_sf() %>% # Convert to sf
+  st_transform(cCRS) %>% # transform to robinson
+  st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
+  rename(FishingCost = layer)
+
+PUs <- PUs %>% 
+  mutate(Fishing = FishingCost$FishingCost)
+
+LogFishing <- log10(FishingCost$FishingCost)
+hist(LogFishing)
+PUs <- PUs %>% 
+  mutate(LogFishing = LogFishing)
+
+#' Plotting log of fishing cost
+
+ggplot() + 
+  geom_sf(data = world) +
+  geom_sf(data = PUs, aes(color=LogFishing)) + 
+  ggtitle("Fishing Cost (Log10)") + 
+  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE)
+
+#(ggCost <- fSpatPlan_PlotCost(Cost, world)) # Plot cost
+
+#FishingCost <- fSpatPlan_Get_Cost(PUs, cCRS) ##make sure cost function is set to fetch .grd in proper directory ##cannot open .grd despite changing pathway in fx
+
+```
 
@@
 
-Overlay shipping raster with polygons to get average intensity of shipping for 1 PU.
+Overlay shipping raster with polygons to get average intensity of shipping for 1 PU. **Check code, no errors but many issues.**
 
-```{r Shipping plot, echo = TRUE}
+```
 
@@
 
-## Fishing
+# Prioritizr problems
 
-```{r Fishing cost, echo = TRUE, warning = FALSE}
+Trial runs
+```{r Prioritizr problems, echo = FALSE, message = FALSE, warning = FALSE}
 
-#adding modified fx here to ensure it uses the right path
+PUs$Fishing[PUs$Fishing <  0.0001] <-  0.0001 #to avoid Warning in presolve_check.OptimizationProblem(compile(x)) :
+# planning units with very high (> 1e+6) or very low (< 1e-6) non-zero cost values note this may be a false positive
 
-fSpatPlan_Get_Cost <- function(PUs, cCRS){
+#' IBA & Fishing cost
 
-  # TODO Add some filtering ability to get the cost only for specific groups
+p_IBA_Fishing <- problem(PUs, features = "IBA", cost_column = "Fishing") %>% 
+  add_min_set_objective() %>%
+  add_relative_targets(0.9) %>%
+  add_binary_decisions() %>%
+  add_gurobi_solver(gap = 0.1, verbose = FALSE)
 
-  FishingCost <- terra::rast("Data/Cost/Cost_Raster_Sum.grd") %>% 
-    terra::as.polygons(trunc = FALSE, dissolve = FALSE, na.rm=FALSE) %>% # Convert to polygon data
-    st_as_sf() %>% # Convert to sf
-    st_transform(cCRS) %>% # transform to robinson
-    st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
-    rename(FishingCost = layer)
-}
+p_IBA_Fishing_Sol <- solve(p_IBA_Fishing)
 
-FishingCost <- fSpatPlan_Get_Cost(PUs, cCRS)
-PUs <- PUs %>% 
-  mutate(Fishing = FishingCost) #issues: no error but cannot check PU dataframe after this: Error in `[.data.frame`(x, i) : undefined columns selected; probably need to restart session
-hist(FishingCost$FishingCost) 
-#FishingCostPlusOne <- FishingCost$FishingCost + 1 #not needed, min >0 for PU=ABNJ
-hist(log10(FishingCost$FishingCost)) #drop after log10(x)=4, x=10^4=10,000? #no longer needed since in ABNJ max val = ##USD 5000, and most values are 10^1 - 10^2 USD
+p_IBA_Fishing_Sol_Plot <- ggplot() +
+  ggtitle("IBA vs Fishing, Target = 0.9") +
+  geom_sf(data = world, colour ="grey70") +
+  geom_sf(data = p_IBA_Fishing_Sol, aes(fill = as.factor(solution_1)), size = 0.05) +
+  scale_fill_manual(values = c("NA", "red")) +
+  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE) +
+  theme_bw() +
+  #scale_colour_gradientn(colours = colorspace::diverge_hcl(7)) +
+  theme(legend.title = element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank())
 
-ggplot() + 
-  geom_sf(data = world) +
-  geom_sf(data = FishingCost, aes(color=FishingCost)) +
-ggtitle("Fishing Cost All Values") + 
-  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
+p_IBA_Fishing_Sol_Plot #looks correct
 
+#' Active Vents & Fishing cost & target
+
+p_Active_Vents_Fishing <- problem(PUs, features = "Active_Vents", cost_column = "Fishing") %>% 
+add_min_set_objective() %>%
+  add_relative_targets(0.7) %>%
+  add_binary_decisions() %>%
+  add_gurobi_solver(gap = 0.1, verbose = FALSE)
+
+p_Active_Vents_Fishing_Sol <- solve(p_Active_Vents_Fishing)
+
+p_Active_Vents_Fishing_Sol_Plot <- ggplot() +
+  ggtitle("Active Vents vs Fishing, Targets = 0.7") +
+  geom_sf(data = world, colour ="grey80") +
+  geom_sf(data = p_Active_Vents_Fishing_Sol, aes(fill = as.factor(solution_1)), size = 0.05) +
+  scale_fill_manual(values = c("NA", "red")) +
+  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE) +
+  theme_bw() +
+  #scale_colour_gradientn(colours = colorspace::diverge_hcl(7)) +
+  theme(legend.title = element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank())
+
+p_Active_Vents_Fishing_Sol_Plot # also looks correct
+```
+
+Full run
+
+```{r Prioritizr problems2, echo = FALSE, message = FALSE, warning = FALSE}
+#' All features & Fishing cost & targets
+
+p_ConsFeat_Fishing <- problem(PUs, features = c("IBA", "IMMA", "NEIO_09", "NWIO_14", "SIO_11", "SIO_19", "SIO_22", "SIO_23", "SIO_30", "SIO_32", "SIO_35", "SIO_36", "SIO_37","Seamounts", "Plateaus", "Active_Vents", "Inactive_Vents"), cost_column = "Fishing") %>%
+  add_min_set_objective() %>%
+  add_relative_targets(c(0.3, 0.3, 0.3, 0.3, 0.3, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.3, 0.3, 0.7, 0.7)) %>%
+  add_binary_decisions() %>%
+  add_gurobi_solver(gap = 0.1, verbose = FALSE)
+
+p_ConsFeat_Fishing_Sol <- solve(p_ConsFeat_Fishing)
+#plot(p_ConsFeat_Fishing_Sol, main = "solution")
+#lines(p_ConsFeat_Fishing_Sol[p_ConsFeat_Fishing_Sol$solution_1 > 0.5, ], col = "darkgreen", lwd = 2)
+
+p_ConsFeat_Fishing_Sol_Plot <- ggplot() +
+  ggtitle("Cons Feats vs Fishing, 0.3<Targets<0.7") +
+  geom_sf(data = world, colour ="grey80") +
+  geom_sf(data = p_ConsFeat_Fishing_Sol, aes(fill = as.factor(solution_1)), size = 0.05) +
+  scale_fill_manual(values = c("NA", "red")) +
+  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE) +
+  theme_bw() +
+  #scale_colour_gradientn(colours = colorspace::diverge_hcl(7)) +
+  theme(legend.title = element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank())
+
+p_ConsFeat_Fishing_Sol_Plot # shows that fishing cost is reduced between 40°-60°S, but maybe would be good to reduce target for IBAs/IMMAs or assign Longhurst to distribute better across latitudes
+
+```
+
+Run with lock out: no errors in  .R but won't knit
+
+```
+
+#' All features & Fishing cost & targets & locked-out exploration + reserved areas
+
+PUs["locked_out"] = PUs$ExplPMN | PUs$ExplPMS | PUs$ResPMN # create locked_out column (logical).
+If =<1 is TRUE, then locked_out is TRUE # https://stackoverflow.com/questions/54507486/merging-two-true-false-dataframe-columns-keeping-only-true
+
+#PUs["locked_out"] <- paste(PUs$ExplPMN, PUs$ExplPMS, PUs$ResPMN)
+#PUs %>% mutate(sum = rowSums(across(where(is.logical))))rowSums(Y) > 0
+#PUs$LockedOut <- ifelse(PUs$ExplPMN == "TRUE" | PUs$ExplPMS == "TRUE" | PUs$ResPMN == "TRUE")
+
+p_ConsFeat_Fishing_LockedOut <- problem(PUs, features = c("IBA", "IMMA", "NEIO_09", "NWIO_14", "SIO_11", "SIO_19", "SIO_22", "SIO_23", "SIO_30", "SIO_32", "SIO_35", "SIO_36", "SIO_37","Seamounts", "Plateaus", "Active_Vents", "Inactive_Vents"), cost_column = "Fishing") %>%
+  add_min_set_objective() %>%
+  add_relative_targets(c(0.3, 0.3, 0.3, 0.3, 0.3, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.3, 0.3, 0.7, 0.7)) %>%
+  add_binary_decisions() %>%
+  add_locked_out_constraints(locked_out = "locked_out") %>%
+  add_gurobi_solver(gap = 0.1, verbose = FALSE)
+
+p_ConsFeat_Fishing_LockedOut_Sol <- solve(p_ConsFeat_Fishing_LockedOut)
+
+p_ConsFeat_Fishing_LockedOut_Sol_Plot <- ggplot() +
+  ggtitle("Cons Feats vs Fishing, Mining Contractors Locked Out, 0.3<Targets<0.7") +
+  geom_sf(data = world, colour ="grey80") +
+  geom_sf(data = p_ConsFeat_Fishing_Sol, aes(fill = as.factor(solution_1)), size = 0.05) +
+  scale_fill_manual(values = c("NA", "red")) +
+  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE) +
+  theme_bw() +
+  #scale_colour_gradientn(colours = colorspace::diverge_hcl(7)) +
+  theme(legend.title = element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank())
+
+p_ConsFeat_Fishing_LockedOut_Sol_Plot
 ```
@@
 
-#section not needed anymore since no outliers when PUs=ABNJ
+#' All features & Fishing cost & targets & locked-out exploration + reserved areas, change targets to check solver works
 
-FishingCostOutliers <- filter(FishingCost, Cost >= 22000) # check outliers
+length(PUs$Plateaus[PUs$Plateaus==TRUE]) # check count of PUs that have plateaus inside
 
-hist(FishingCostOutliers$Cost)
+p_check_targets <- problem(PUs, features = c("IBA", "IMMA", "NEIO_09", "NWIO_14", "SIO_11", "SIO_19", "SIO_22", "SIO_23", "SIO_30", "SIO_32", "SIO_35", "SIO_36", "SIO_37","Seamounts", "Plateaus", "Active_Vents", "Inactive_Vents"), cost_column = "Fishing") %>%
+add_min_set_objective() %>%
+  add_relative_targets(c(0.001, 0.001, 0.03, 0.03, 0.03, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.99, 0.99, 0.99, 0.99)) %>%
+  add_binary_decisions() %>%
+  add_locked_out_constraints(locked_out = "locked_out") %>%
+  add_gurobi_solver(gap = 0.1, verbose = FALSE)
 
-ggplot() + 
-  geom_sf(data = world) +
-  geom_sf(data = FishingCostOutliers, aes(color=Cost)) +
-ggtitle("Fishing Cost Outliers (>50,000 USD)") + 
-  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
+p_check_targets_sol <- solve(p_check_targets)
 
-FishingCostNoOutliers <- filter(FishingCost, Cost <= 22000) # remove outliers
+p_check_targets_sol_plot <- ggplot() +
+  ggtitle("0.001<Targets<0.9") +
+  geom_sf(data = world, colour ="grey80") +
+  geom_sf(data = p_ConsFeat_Fishing_Sol, aes(fill = as.factor(solution_1)), size = 0.05) +
+  scale_fill_manual(values = c("NA", "red")) +
+  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE) +
+  theme_bw() +
+  #scale_colour_gradientn(colours = colorspace::diverge_hcl(7)) +
+  theme(legend.title = element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank())
 
-hist(FishingCostNoOutliers$Cost) #(49256 obs. = 524 outliers removed when #cut-off = 20,000)
+p_check_targets_sol_plot #correct
 
-ggplot() + 
-  geom_sf(data = world) +
-  geom_sf(data = FishingCostNoOutliers, aes(color=Cost)) +
-ggtitle("Fishing Cost Without Outliers") + 
-  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
-
-```
-
-# Prioritizr problems
-
-## Basic run (Cons features)
-
-*Code based on Alvise's work, not as code here to avoid issues when knitting*
-
-```
-
-p_ConsFeats <- problem(PUs, features = Cons_feats$Features) %>% #issues because no cost_column
-  add_min_set_objective() %>%
-  add_relative_targets(Cons_feats$Targets) %>% 
-  add_binary_decisions() %>%
-  add_lpsymphony_solver(verbose = FALSE)
-
-```
-
-### Cons features, fishing
-
-*code based on Alvise's work, not as code here to avoid issues when knitting*
-
-```
-
-p_ConsFeat_Fishing <- problem(PUs, features = Cons_feats$Features, cost_column = PUs$FishingCost) %>% 
-  add_min_set_objective() %>% 
-  add_relative_targets(Cons_feats$Targets) %>% 
-  add_binary_decisions() %>%
-  add_lpsymphony_solver(verbose = FALSE)
-
-```
-
-### Cons features, mining
-
-*code based on Alvise's work, not as code here to avoid issues when knitting*
-
-*PMNCost will probably be created based on Sharma (2011) & CRU Consulting (2019)?*
-
-[prioritizr.net](https://prioritizr.net/reference/add_locked_out_constraints.html) says to use structure below to lock out polygons but need to look more into it
-
-# create problem with added locked out constraints using spatial polygon data
-
-```
-use:
-
-locked_out <- sim_pu_polygons[sim_pu_polygons$locked_out == 1, ]
-p5 <- p1 %>% add_locked_out_constraints(locked_out)
-```
-or try
-
-```
-
-p_ConsFeat_Mining <- problem(PUs, features = Cons_feats$Features, cost_column = PUs$PMNCost) %>% 
-  add_min_set_objective() %>% 
-  add_relative_targets(Cons_feats$Targets) %>% 
-  add_binary_decisions() %>%
-  add_lpsymphony_solver(verbose = FALSE) %>%
-  add_locked_out_constraints("Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
-```
-
-### Cons features, mining + fishing
-
-*Assumes targets (ConsFeatures$amount) are set, code based on Alvise's work, not as code here to avoid issues when knitting*
-
-```
-
-p_ConsFeat_MiningFishing <- problem(PUs, features = Cons_feats, cost_column = PUs$MiningFishing) %>% 
-  add_min_set_objective() %>% 
-  add_relative_targets(Cons_features$Targets) %>%
-  add_binary_decisions() %>%
-  add_lpsymphony_solver(verbose = FALSE) %>%
-  add_locked_out_constraints("Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
-```
-
-### Cons features, shipping
-
-*Not sure what kind of data transformation is needed to have a file with present/absent shipping routes*
-
-*Or we can try estimating cost of shipping (question at start of .rmd) and combining with other costs into 1 layer*
-
-*here, Shipping = mean values of pixels overlaid on PUs, so intensity of shippings; not USD*
-
-```{r Shipping problem, echo = TRUE, warning = FALSE}
-
-p_ConsFeat_Shipping <- problem(PUs, features = Cons_feats$Features, cost_column = "PUs$Shipping") %>% 
-  add_min_set_objective() %>% 
-  add_relative_targets(Cons_feats$Targets) %>% 
-  add_binary_decisions() %>%
-  add_lpsymphony_solver(verbose = FALSE) #%>%
-  #add_locked_out_constraints("Shipping","Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
 ```

### 2022-03-09 SpatPlan_Master_Lea_v2.R
with prio solved and plotted
same as rmd

@@ -1,16 +1,12 @@
 #' ---
 #' title: "Create a Spatial Plan"
-#' author: "Jason D. Everett"
-#' affiliation: "UQ/CSIRO/UNSW"
-#' date: "Last compiled on `r format(Sys.time(), '%A %d %B %Y')`"
+#' author: "Jason D. Everett, modified by Lea Fourchault"
+#' affiliation: "UQ/CSIRO/UNSW/Tropimundo"
+#' date: "Last compiled on March 8th, 2022"
 #' output: 
 #'   html_document
 #' ---
 #'
-
-##+ setup, include=FALSE
-knitr::opts_chunk$set(warning=FALSE, cache=FALSE, message=FALSE)
-# knitr::opts_chunk$set(collapse = TRUE, comment = "", warning = "off")
 
 #' ## Overview
 #' This code has been written to simplify the process for running a _prioritizr_ analysis on a given region. It is still a work in progress so feel free to submit pull requests with new features and code improvements.
@@ -61,7 +57,9 @@ fSpatPlan_Get_Polyg <- function(filen, PUs){
 cCRS <- "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs" # Robinson projection
 world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
   st_transform(cCRS)
-Region <- c(xmin = 18, xmax = 130, ymin = -39, ymax = 34) # Indian Ocean limits
+#Region <- c(xmin = 18, xmax = 140, ymin = -39, ymax = 34) 
+Region <- c(xmin = 20, xmax = 120, ymin = -60, ymax = 20) 
+# Indian Ocean limits: CIA World Factbook (but see: According to the International Hydrographic Organization (IHO) and the United Nations Oceans Atlas, the area from 40 degrees S latitude to 60 degrees S latitude is included in the Indian Ocean. The area that encircles the globe from 60 degrees S latitude to the coast of Antarctica is called The Great Southern Ocean. The Indian Ocean's width extends from 45 degrees E longitude to 110 degrees East longitude.) 
[…]

### 2022-03-09 SpatPlan_Master_Lea_v2.R
adding Alvise's function
functions for comparing solutions & kappa correlation, need to modify for my project

@@ -253,6 +253,7 @@ nr <- st_nearest_feature(PUs, longhRob)
 PUs <- PUs %>% 
   mutate(ProvCode = longh$ProvCode[nr],
          ProvDescr = longh$ProvDescr[nr])
+PUs$ProvCode
 
 ggplot() + 
   geom_sf(data = world, color = "grey20", fill = "grey20", size = 0.1, show.legend = FALSE) +
@@ -603,3 +604,172 @@ p_check_targets_sol_plot <- ggplot() +
         axis.title.y = element_blank())
 
 p_check_targets_sol_plot
+
+
+#' ## Comparing solutions
+
+# based on Alvise's functions:
+
+fCompareSolutions <- function(ctrl, rmv){
+  
+  #Select only the ID column and the solution column; change the name of the columns in 'cellID' and 'ctrl_sol'  
+  ctrl2 <- ctrl %>%
+    dplyr::select(ID, solution_1) %>%
+    rename(cellID = ID, ctrl_sol = solution_1)
+  
+  rmv2 <- rmv %>%
+    dplyr::select(ID, solution_1) %>%
+    rename(cellID = ID, rmv_sol = solution_1)
+  
+  #Create an object with the two solutions
+  soln <- ctrl2
+  soln$rmv_sol <-  rmv2$rmv_sol
+  
+  #Combine the result and call them differently if a PUs in present in both the solution, or only in one of them
+  soln <- mutate(soln, Combined = as.numeric(ctrl_sol + rmv_sol)) %>%
+    mutate(Compare = case_when(Combined == 2 ~ "Same",
+                               ctrl_sol == 1 & rmv_sol == 0 ~ "Removed (-)",
+                               ctrl_sol == 0 & rmv_sol == 1 ~ "Added (+)",
+                               Combined == 0 ~ "Not Selected"),
+           Compare = factor(Compare, levels = c("Added (+)", "Same", "Removed (-)", "Not Selected"))) %>%
+    filter(!is.na(Compare))
+  
+  # label_loc$label <- paste0("Area: ",round((sum(rmv$solution_1) - sum(ctrl$solution_1))/ sum(ctrl$solution_1)*100, 1), "%\nCost: ",round((sum(rmv$solution_1*rmv$cost) - sum(ctrl$solution_1*ctrl$cost))/ sum(ctrl$solution_1*ctrl$cost)*100, 1), "%\n")
[…]
